---
title: "Poudre Water Quality Network Data Availability Report"
author: "Maggie Rouin"
date: "`r Sys.Date()`"
output: html_document
---

# set up

## load packages
```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.dim = c(12,10))
# loading packages
package_loader <- function(x) {
  if (x %in% installed.packages()) {
    suppressMessages({
      library(x, character.only = TRUE)
    })
  } else {
    suppressMessages({
      install.packages(x)
      library(x, character.only = TRUE)
    })
  }
}

invisible(
  lapply(c("arrow",
           "data.table",
           "httr",
           "httr2",
           "tidyverse",
           "lubridate",
           "zoo",
           "padr",
           "stats",
           "RcppRoll",
           "yaml",
           "ross.wq.tools",
           "here",
           "furrr", 
           "dataRetrieval", 
           "cdssr", 
           "ggplot2", 
           "ggthemes", 
           "patchwork",
           "sf", 
           "plotly",
           "viridisLite",
           "tmap",
           "knitr",
           "flextable",
           "DT"
  ),
  package_loader)
)

`%nin%` = Negate(`%in%`)
```


## sites, parameters, and plots
```{r, include = FALSE}
if ({
  require(systemfonts)
  ("Roboto" %in% system_fonts()$family)
}) {
  ROSS_theme <- theme_bw() +
    theme(plot.title = element_text(hjust = 0.5, face = 'bold', family = "Roboto"),
          plot.subtitle = element_text(hjust = 0.5, family = "Roboto"))
} else {
  message("You do not have the Roboto font family installed on your computer, currenly using ggplot default text family.
          See ROSS_themes.R for directions to install the font family on your computer.")
  ROSS_theme <- theme_bw() +
    theme(plot.title = element_text(hjust = 0.5, face = 'bold'),
          plot.subtitle = element_text(hjust = 0.5))
}


#Site names
site_names <- tibble(site = c("pbd", "bellvue", "salyer", "udall", "riverbend", "cottonwood" ,"elc",  "archery", "riverbluffs"), 
                     site_name = c("Canyon Mouth", "Tamasag", "Legacy", "Lincoln", "Timberline", "Prospect" ,"Boxelder",  "Archery", "River Bluffs"), 
                     site_color = c( "#000000" , "#185BB4","#009DD1","#00F5DC", "#00F0C0", "#00D684", "#00BD48",  "#00DB58", "#0FFF6B" ) )

#Site Order from up to downstream

site_order = c("Canyon Mouth", "Bellvue", "Salyer", "Udall", "Riverbend", "Cottonwood", "ELC", "Archery", "River Bluffs")


#ROSS Colors

ROSS_lt_pal <- c("#002EA3", "#E70870", "#256BF5",
                 "#745CFB", "#1E4D2B", "#56104E")

ROSS_dk_pal <- c("#1E4D2B", "#E70870", "#256BF5",
                 "#FFCA3A", "#745CFB", "#C3F73A")

ROSS_acc_pal <- c("#56104E", "#745CFB", "#FFCA3A", "#1E4D2B")

# Parameter labels
labels <- tibble(param = c("Turbidity", "Specific Conductivity", "Depth", "Chl-a Fluorescence", "FDOM Fluorescence", "Temperature", "DO", 'pH', "Flow"),
                 label = c("Turbidity (NTU)", "Specific Conductivity (uS/cm)", "Depth (ft)", "Chl-a Fluorescence (RFU)", "FDOM Fluorescence (RFU)", "Temperature (C)", "DO (mg/L)", 'pH', "Flow (cfs)"))


chem_expressions <- list(
  "Turbidity" = expression(Turbidity ~ (NTU)),
  "Chl-a Fluoresence"      = expression(Chl-a ~ Fluoresence ~ (RFU)),
  "FDOM Fluoresence"      = expression(FDOM ~ Fluoresence ~ (RFU)),
  "Specific Conductivity" = expression(Specific ~ Conductivity ~ (mu ~ S ~ cm^{-1})),
  "Depth"      = expression(Depth ~ (ft)),
  "Temperature" = expression(Temperature ~ (degree ~ C)),
  "DO"        = expression(DO ~ (mg ~ L^{-1})),
  "Flow"      = expression(Flow ~ (cfs)),
  "pH"        = expression(pH)
)


params <- paste(labels$param, collapse = "|")

```


## read in data
```{r, include = FALSE}
all_data <- list.files(
  "operations/reporting/pwqn_data_report/data/hydroshare_pub",
  pattern = "^CSU_ROSS_\\d{4}_cleaned\\.csv$",
  full.names = TRUE,
  recursive = TRUE) |> 
  map(~ read_csv(.x, show_col_types = FALSE)) |> 
  list_rbind()

pwqn_data <- all_data |>
  filter(value_final != "NA") |> 
  mutate(DateTime_UTC = with_tz(DateTime_UTC, tz = "MST")) |> 
  rename(DateTime_MT = DateTime_UTC) |> 
  mutate(
    year = year(DateTime_MT),
    month = month(DateTime_MT),
    year_month = floor_date(DateTime_MT, unit = "month")) |> 
  left_join(labels, by = c( "parameter" = "param")) |> 
  left_join(site_names, by = "site") |> 
  filter(!is.na(site_name))

pwqn_data_clean <- pwqn_data |>
  distinct(DateTime_MT, site_name, parameter, .keep_all = TRUE)

print(paste("Original records:", nrow(pwqn_data)))
print(paste("Records after removing duplicates:", nrow(pwqn_data_clean)))
print(paste("Duplicates removed:", nrow(pwqn_data) - nrow(pwqn_data_clean)))

check_duplicates <- pwqn_data_clean |>
  group_by(DateTime_MT, site_name, parameter) |>
  filter(n() > 1)

print(paste("Duplicates remaining:", nrow(check_duplicates)))

counts <- pwqn_data_clean |> 
  count(site = "archery", parameter = "pH", month = 2)

```


# Graphs
```{r echo=FALSE, fig.width=30, fig.height=30, out.width="100%"}
site_order <- site_names$site_name

#yearly
yearly_percent <- pwqn_data_clean |> 
  left_join(site_names, by = "site_name") |>
  group_by(year, parameter, site_name) |>
  summarise(actual_records = n(),
    .groups = 'drop') |> 
  mutate(expected_records = if_else(
      leap_year(year),
      366 * 24 * 60 / 15,
      365 * 24 * 60 / 15),
    data_pct = (actual_records / expected_records) * 100,
    site_name = factor(site_name, levels = site_order)) |> 
  arrange(year)

yearly_plot <- yearly_percent |> 
  ggplot(aes(x = factor(year), y = data_pct, fill = site_name)) +
  geom_col(position = "dodge") +
  facet_wrap(~parameter, scales = "free_y", ncol = 2) +
  scale_fill_manual(
    name = "Site",
    values = setNames(site_names$site_color, site_names$site_name)
  ) +
  labs(title = "Data Availability by Year",
    x = "Year",
    y = "Availability (%)",
    fill = "Site") +
  ylim(0, 105)

yearly_plot

#monthly
monthly_percent <- pwqn_data_clean |> 
  mutate(month = month(DateTime_MT),
         year = year(DateTime_MT)) |> 
  group_by(year, month, parameter, site_name) |>
  summarise(actual_records = n(),
    .groups = 'drop') |> 
  mutate(days_in_month = case_when(
      month == 2 ~ if_else(year %in% c(2020, 2024), 29, 28),
      month %in% c(4, 6, 9, 11) ~ 30, TRUE ~ 31),
    expected_records = days_in_month * 24 * 60 / 15,
    data_pct = (actual_records / expected_records) * 100) |>
  left_join(site_names, by = "site_name") |>
  mutate(month_year = ym(paste(year, month, sep = "-")),
    site_name = factor(site_name, levels = site_order)) |> 
  arrange(month, month_year, parameter, site_name)

monthly_plot <- monthly_percent |> 
  ggplot(aes(x = month_year, y = data_pct, fill = site_name)) +
  geom_col(position = "dodge") +
  facet_wrap(~parameter, scales = "free_y", ncol = 2) +
  scale_fill_manual(name = "Site",
    values = setNames(site_names$site_color, site_names$site_name)) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "4 months") +
  labs(title = "Monthly Data Availability",
    x = "Month-Year",
    y = "Availability (%)",
    fill = "Site") +
  ylim(0, 105) +
  theme(axis.text.x = element_text(angle = 90))

monthly_plot
```
These plots depict the yearly and monthly data availability percentages for each parameter across all sites. As the years progress, the data availability increases across all parameters.

# Heatmaps
```{r echo=FALSE, fig.width=30, fig.height=50, out.width="100%"}
#all data heatmap
combined_monthly <- monthly_percent |> 
  mutate(fake_date = ymd(paste(year, month, "01", sep = "-"))) |> 
  ggplot(aes(x = fake_date, y = site_name, fill = data_pct)) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_gradient(low = "red", high = "green", limits = c(0, 100)) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months") +
  facet_wrap(~parameter, scales = "free_y", ncol = 1) +
  labs(title = "Data Availability",
       x = "Month",
       y = "Site",
       fill = "Availability (%)")

print(combined_monthly)

```
This heatmap depicts the data availability at each site across the data collection timeframe (March 2019- November 2025). Yearly data availability is highest in 2024 and 2025 and monthly data availability is highest in the late summer months. 

# Timeseries
```{r echo=FALSE, fig.width=30, fig.height=30, out.width="100%"}
#monthly boxplots
param_units <- pwqn_data_clean |>
  select(parameter, units) |>
  distinct() |>
  mutate(axis_label = paste0(parameter, " (", units, ")")) |>
  pull(axis_label, name = parameter)

facet_labeler <- labeller(parameter = param_units)

log_params <- c("Turbidity", "Specific Conductivity", "Chl-a Fluorescence")

pwqn_data_clean <- pwqn_data_clean |>
  mutate(site_name = factor(site_name, levels = site_order))

p_log <- pwqn_data_clean |>
  filter(parameter %in% log_params, value_final > 0) |>
  ggplot(aes(x = factor(year), y = value_final, fill = site_name)) +
  geom_boxplot(size = 0.25, width = 0.6, position = position_dodge(width = 0.75), outlier.size = 0.5) +
  facet_wrap(~parameter, scales = "free_y", labeller = facet_labeler, ncol = 1) +
  scale_y_log10() +
  scale_fill_manual(values = colorRampPalette(c("darkgreen", "lightblue"))(n_distinct(pwqn_data_clean$site_name))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
        axis.title.x = element_text(margin = margin(t = 15)), axis.title.y = element_blank()) +
  labs(x = "Year",
       fill = "Site Name")

p_linear <- pwqn_data_clean |>
  filter(!parameter %in% log_params) |>
  ggplot(aes(x = factor(year), y = value_final, fill = site_name)) +
  geom_boxplot(size = 0.25, width = 0.6, position = position_dodge(width = 0.75), outlier.size = 0.5) +
  facet_wrap(~parameter, scales = "free_y", labeller = facet_labeler, ncol = 1) +
  scale_fill_manual(values = colorRampPalette(c("darkgreen", "lightblue"))(n_distinct(pwqn_data_clean$site_name))) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank(), 
        legend.title = element_blank(), legend.position = "none") +
  labs(title = "Yearly Parameter Averages by Site (upstream to downstream)")

combined_plot <- (p_linear / p_log) +
  plot_layout(heights = c(2, 1), guides = "collect") +
  theme(legend.position = "right")

combined_plot <- wrap_elements(combined_plot) +
  labs(tag = "Value") +
  theme(plot.tag = element_text(size = 11, angle = 90),
        plot.tag.position = "left")

combined_plot
```
This boxplot depicts the average values of each parameter across the sites in the PWQN, from upstream to downstream. The boxplot indicates that averages for pH, temperature, chl-a flourescence, and specific conductivity increase as you move downstream in the network. There is also a slight decrease in dissolved oxygen averages in the downstream sites.


# Historic Range
```{r echo=FALSE, fig.width=30, fig.height=30, out.width="100%"}
#bands
pct_bands <- list(
  list(ymin = "p0",  ymax = "p5",   label = "Extreme (<5% / >95%)",              color = "#8E857A"),
  list(ymin = "p5",  ymax = "p10",  label = "Much below/above (5-10% / 90-95%)", color = "#E8DAB2"),
  list(ymin = "p10", ymax = "p25",  label = "Below/above normal (10-25% / 75-90%)", color = "#F0F7EB"),
  list(ymin = "p25", ymax = "p75",  label = "Normal range (25-75%)",             color = "#9CC4DE"),
  list(ymin = "p75", ymax = "p90",  label = "Below/above normal (10-25% / 75-90%)", color = "#F0F7EB"),
  list(ymin = "p90", ymax = "p95",  label = "Much below/above (5-10% / 90-95%)", color = "#E8DAB2"),
  list(ymin = "p95", ymax = "p100", label = "Extreme (<5% / >95%)",              color = "#8E857A")
)

band_colors <- setNames(
  vapply(pct_bands, `[[`, "", "color"),
  vapply(pct_bands, `[[`, "", "label")
)[!duplicated(vapply(pct_bands, `[[`, "", "label"))]

#daily summary function
daily_summary <- function(x) {
  q <- quantile(x, probs = c(0, .05, .10, .25, .75, .90, .95, 1), na.rm = TRUE)
  tibble(p0 = q[1], p5 = q[2], p10 = q[3], p25 = q[4],
         p75 = q[5], p90 = q[6], p95 = q[7], p100 = q[8],
         Mean = mean(x, na.rm = TRUE))
}

#remove legend duplicates
updated_legend <- function(p) {
  seen <- character(0)
  for (i in seq_along(p$x$data)) {
    nm <- p$x$data[[i]]$name
    if (is.null(nm)) next
    
    # strip wrapping "(" ")" and trailing ",<number>" that ggplotly adds per facet panel
    clean_nm <- sub("^\\((.*),\\d+\\)$", "\\1", nm)
    
    if (clean_nm %in% seen) {
      p$x$data[[i]]$showlegend <- FALSE
    } else {
      seen <- c(seen, clean_nm)
      p$x$data[[i]]$name <- clean_nm   # restore clean label for the one that's shown
    }
  }
  p
}

#temps plot
historic_temps_all_sites <- function(data = pwqn_data_clean,
                                      plot_title = "Historical Water Temperature by Site",
                                      sites = c("Canyon Mouth", "Tamasag", "Legacy", "Lincoln",
                                                "Timberline", "Prospect", "Boxelder", "Archery", "River Bluffs"),
                                      incl_winter = FALSE,
                                      temp_param_name = "Temperature") {

  temp_df <- data |>
    filter((site_name %in% sites | site %in% sites),
           grepl(temp_param_name, parameter, ignore.case = TRUE))
  if (nrow(temp_df) == 0) stop("No temperature data found for the specified sites.")

  temp_data <- temp_df |>
    mutate(datetime = as.Date(DateTime_MT),
           month_day = format(datetime, "%m-%d"),
           value = value_final) |>
    filter(!is.na(value))
  if (!incl_winter) temp_data <- filter(temp_data, !month(datetime) %in% c(12, 1, 2))

  ref_year <- 2000
  daily_stats <- temp_data |>
    group_by(site_name, month_day) |>
    reframe(daily_summary(value)) |>
    mutate(plot_date = as.Date(paste0(ref_year, "-", month_day), "%Y-%m-%d")) |>
    filter(!is.na(plot_date)) |>
    arrange(site_name, plot_date) |>
    mutate(site_name = factor(site_name, levels = sites))

  ribbon_layers <- lapply(pct_bands, function(b) {
    geom_ribbon(aes(ymin = .data[[b$ymin]], ymax = .data[[b$ymax]], fill = b$label))
  })

  p <- ggplot(daily_stats, aes(x = plot_date)) +
    ribbon_layers +
    geom_line(aes(y = Mean, color = "Daily Average (Mean)"), linewidth = 0.5) +
    facet_wrap(~site_name, ncol = 1, scales = "fixed") +
    scale_fill_manual(name = "Percentile Range", values = band_colors, breaks = names(band_colors)) +
    scale_color_manual(name = "", values = c("Daily Average (Mean)" = "#222222")) +
    scale_x_date(date_labels = "%b", date_breaks = "2 months") +
    labs(title = plot_title, x = "", y = "Temperature (°C)") +
    theme_bw() +
    theme(legend.position = "bottom",
          strip.background = element_rect(fill = "grey90"),
          strip.text = element_text(face = "bold"),
          axis.text.x = element_text(angle = 90, hjust = 1))

  gp <- ggplotly(p)
  updated_legend(gp)
}

all_sites_temp_plot2 <- historic_temps_all_sites(pwqn_data_clean)
all_sites_temp_plot2

```
The timeseries plot shows the historical water temperatures at each site across the year.

# Interactive Tables
```{r echo=FALSE, fig.width=30, fig.height=30, out.width="100%"}
#yearly
format_availability_dt <- function(data, col_names, align_vec) {
  n_cols <- ncol(data)
  center_cols <- which(align_vec == "c") - 1
  datatable(
    data,
    colnames = col_names,
    rownames = FALSE,
    filter = "top",
    extensions = "Buttons",
    options = list(
      pageLength = 25,
      lengthMenu = c(10, 25, 50, 100),
      dom = 'Blfrtip',
      buttons = c('copy', 'csv', 'excel'),
      columnDefs = list(
        list(className = 'dt-center', targets = center_cols)),
      initComplete = JS("function(settings, json) {",
        "$(this.api().table().header()).css({'background-color': 'darkblue', 'color': 'white', 
        'font-weight': 'bold'});","}")),
    class = "stripe hover compact")
}

yearly_data_table_site <- yearly_percent |>
  group_by(year, parameter, site_name) |>
  summarise(
    avg_availability = round(mean(data_pct, na.rm = TRUE), 2),
    .groups = 'drop') |>
  mutate(year = sprintf("%d", year))

yearly_site_dt <- format_availability_dt(
  yearly_data_table_site,
  col_names = c("Year", "Parameter", "Site", "Average Availability (%)"),
  align_vec = c("l", "l", "l", "c"))

yearly_site_dt

yearly_data_table <- yearly_percent |>
  group_by(year, parameter) |>
  summarise(
    avg_availability = round(mean(data_pct, na.rm = TRUE), 2),
    .groups = 'drop') |>
  mutate(year = sprintf("%d", year))

yearly_dt <- format_availability_dt(
  yearly_data_table,
  col_names = c("Year", "Parameter", "Average Availability (%)"),
  align_vec = c("l", "l", "c"))

yearly_dt

# monthly
monthly_table_site <- monthly_percent |>
  mutate(month_name = factor(month.name[month], levels = month.name)) |> 
  group_by(month_name, parameter, site_name) |>
  summarise(
    avg_availability = round(mean(data_pct, na.rm = TRUE), 2),
    .groups = 'drop')

monthly_site_dt <- format_availability_dt(
  monthly_table_site,
  col_names = c("Month", "Parameter", "Site", "Average Availability (%)"),
  align_vec = c("l", "l", "l", "c"))

monthly_site_dt

monthly_data_table <- monthly_percent |>
  mutate(month_name = factor(month.name[month], levels = month.name)) |> 
  group_by(month_name, parameter) |>
  summarise(
    avg_availability = round(mean(data_pct, na.rm = TRUE), 2),
    .groups = 'drop')

monthly_dt <- format_availability_dt(
  monthly_data_table,
  col_names = c("Month", "Parameter", "Average Availability (%)"),
  align_vec = c("l", "l", "c"))

monthly_dt
  
```
These interactive tables are broken up by yearly and monthly overall network data availability and by site.
