---
title: "Processing 2023 Data Pull from HydroVu with Auto Back-Calibration and QA/QC Flagging"
author: "Sam Struthers - CSU ROSSyndicate"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

# Script Purpose & Scope

This script is designed to pull in sensor data from HydroVu for the 2023 monitoring season, apply back-calibrations,run through the automated QA/QC flagging process, prep data for manual data verification/drift correction, and then prepare the final dataset for publication in HydroShare.

Steps:

1.  Setup Libraries, Data Directories, Sites and Time Range

2.  Load in all datasets (HydroVu, downloaded logs, field notes) & tidy

3.  Apply back calibrations across all site-parameters

4.  Pass data into calibration correction tool

5.  Load in back-calibrated data and apply auto-QAQC

6.  Pass data into manual verification and drift correction tools

7.  Apply any additional updates (depth and turbidity corrections in this case)

8.  Prep the data for publication

## Setup Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# loading packages
package_loader <- function(x) {
  if (x %in% installed.packages()) {
    suppressMessages({
      library(x, character.only = TRUE)
    })
  } else {
    suppressMessages({
      install.packages(x)
      library(x, character.only = TRUE)
    })
  }
}

invisible(
  lapply(c("arrow",
           "sf",
           "tmap",
           "data.table",
           "httr2",
           "tidyverse",
           "lubridate",
           "zoo",
           "padr",
           "stats",
           "RcppRoll",
           "yaml",
           "here",
           "ross.wq.tools",
           "furrr",
           "plotly"
  ),
  package_loader)
)

# Set up parallel processing
num_workers <- min(availableCores() - 1, 4) # Use at most 4 workers
plan(multisession, workers = num_workers)
set_furr_options <- furrr_options(
  globals = TRUE,
  packages = c("arrow", "data.table", "httr2", "tidyverse","dplyr", "lubridate", "zoo",
               "padr", "stats", "RcppRoll", "yaml", "here", "ross.wq.tools")
)

# suppress scientific notation to ensure consistent formatting
options(scipen = 999)

`%nin%` = Negate(`%in%`)
```

## Setting up the site/time range/parameters for the data pull

We deploy our sensors in March and pull them by the end of November.
Always triple check the sites listed below to remove sites no longer deployed and add sites deployed this year.
Double check the parameters of interest for spelling and inclusion of all needed parameters.

```{r}
#Adjust dates for specific pull
mst_start <- ymd_hms("2023-02-01 00:00:00", tz = "America/Denver")
mst_end <- ymd_hms("2023-12-30 23:59:59", tz = "America/Denver")

sites <- c(
  #Upper sites
  "chd", "sfm", "pbd","joei", "cbri",
  #Lower sites (using all names in case they got mixed up in HydroVu)
  "tamasag","bellvue","legacy","salyer", "lincoln", "udall", "timberline", "riverbend","prospect", "cottonwood", "boxelder", "elc","archery", "riverbluffs", "boxcreek")

#Setting parameters of interest
parameters <- c("Specific Conductivity", "Temperature", "DO", "pH", "ORP", "Depth", #Standard vars
                "Chl-a Fluorescence", "Turbidity", "FDOM Fluorescence") #optical vars
```

## Establishing directory paths.

These are the folders where raw, and flagged data will be saved.
Update base_folder to the current year and log directory as needed

```{r}
base_folder <- here("data","raw", "sensor","manual_data_verification","2023_cycle")
# Make sure base folder exists and create if not
if (!dir.exists(base_folder)) {
  dir.create(base_folder)
}

#folder with downloaded logs
log_directory <- here("data","raw", "sensor",  "log_download", "2023")

#Sub folders for raw and flagged data
staging_directory <- here(base_folder,"hydro_vu_pull" ,"raw_data")
flagged_temp_dir <- here(base_folder,"hydro_vu_pull" ,"flagged_data_temp")
flagged_directory <- here(base_folder,"hydro_vu_pull" ,"flagged_data")

#check to make sure these all exist and create if needed
if (!dir.exists(staging_directory)) {
  dir.create(staging_directory, recursive = TRUE)
}
if (!dir.exists(flagged_temp_dir)) {
  dir.create(flagged_temp_dir, recursive = TRUE)
}
if (!dir.exists(flagged_directory)) {
  dir.create(flagged_directory, recursive = TRUE)
}

```

## Read in thresholds and credentials

Sensor thresholds are from In-Situ Inc spec sheets and seasonal_thresholds are created using `ross.wq.tools::make_threshold_table()` using cleaned data from previous years.
mWater and HydroVu creds are created from their respective websites, contact Sam Struthers, Juan De La Torre or Katie Willi for help obtaining these.

```{r}
# Read in the threshold data first
sensor_thresholds <- read_yaml(here("data","derived","auto_qaqc_files", "thresholds",  "sensor_spec_thresholds.yml"))
season_thresholds <- read_parquet(here("data","derived","auto_qaqc_files", "thresholds", "seasonal_thresholds20260304-T175010Z.parquet"), show_col_types = FALSE) %>%
  fix_site_names()

# Read in credentials
mWater_creds <- read_yaml(here("creds", "mWaterCreds.yml"))

hv_creds <- read_yaml(here("creds", "HydroVuCreds.yml"))
hv_token <- hv_auth(client_id = as.character(hv_creds["client"]),
                    client_secret = as.character(hv_creds["secret"]))

```

# Data Pulls

Once these sections have been run and data have been checked using plots below, these sections should be commented out as they will take a long time to re-run.

## Pulling in older field notes

```{r}
old_field_notes <- readxl::read_excel(here("data","raw","field_notes","sensor_field_notes.xlsx")) %>% 
  mutate(DT = (paste0(date, " ", start_time_mst))) %>%
  mutate(DT = ymd_hm(DT) + hours(7)) %>%
  arrange(DT) %>%
  mutate(DT_round = round_date(DT, "15 minutes")) %>%
  mutate(DT_round = with_tz(DT_round, tzone = "UTC"),
         last_site_visit = with_tz(DT_round, tzone = "UTC"),
         DT_join = as.character(DT_round),
         sonde_moved = NA,
         sonde_employed = case_when(is.na(sensor_deployed) & is.na(sensor_pulled) ~ NA,
                                    sensor_deployed == "x" ~ 0, 
                                    sensor_pulled == "x" ~ 1), 
         site = ifelse(site == "rist", "bellvue", site)) %>%
  ross.wq.tools::fix_site_names()%>%
  #add extra cols
  mutate(rdo_pre_clean = NA, 
         rdo_post_clean = NA,
         rdo_post_cal = NA,
         cond_pre_clean = NA,
         cond_post_clean = NA,
         cond_post_cal = NA,
         ph_pre_clean = NA,
         ph_post_clean = NA,
         ph_post_cal = NA,
         orp_pre_clean = NA,
         orp_post_clean = NA,
         orp_post_cal = NA,
         turb_pre_clean = NA,
         turb_post_clean = NA,
         turb_post_cal = NA,
         fdom_pre_clean = NA,
         fdom_post_clean = NA,
         fdom_post_cal = NA,
         chla_pre_clean = NA,
         chla_post_clean = NA,
         chla_post_cal = NA)%>%
  select(-sensors_cleaned, -wiper_working, -cal_report_collected)
 

```

## Pulling in field notes from mWater

```{r}
# Pull in all notes
mWater_data <- load_mWater(creds = mWater_creds)
# Only extract sensor visits
all_field_notes <- grab_mWater_sensor_notes(mWater_api_data = mWater_data)%>%
  mutate(DT_round = with_tz(DT_round, tzone = "UTC"), # grab_mWater_sensor_notes converts this to MST, converting back to UTC for DS use
         last_site_visit = with_tz(last_site_visit, tzone = "UTC"),
         DT_join = as.character(DT_round))%>%
  bind_rows(old_field_notes)%>%
  ross.wq.tools::fix_site_names()
# Only extract sensor malfunctions
sensor_malfunction_notes <- grab_mWater_malfunction_notes(mWater_api_data = mWater_data)%>%
  mutate(start_DT = with_tz(start_DT, tzone = "UTC"),# grab_mWater_malfunction_notes converts this to MST, converting back to UTC for DS use
         end_DT = with_tz(end_DT, tzone = "UTC"))
```

## Pulling in the sensor data from HydroVu

This section will take quite a while to load (\~ 2-4 min per month of data per site).
Once data has been pulled and checked, comment this section out to avoid re-running it.

```{r}
# # Get all the site info from HydroVu
hv_sites <- hv_locations_all(hv_token)
#trim down to just the sites/types of devices we want
hv_sites <- hv_sites %>%
  #Remove Vulinks
  filter(!grepl("vulink", name, ignore.case = TRUE))%>%
  #keep 2023 sites or chd
  filter(grepl("2023", name, ignore.case = TRUE) | grepl("chd", name, ignore.case = T))


walk(sites,
     function(site) {
       message("Requesting HV data for: ", site)
       api_puller(
         site = site,
         start_dt = with_tz(mst_start, tzone = "UTC"),
         end_dt = with_tz(mst_end, tzone = "UTC"),
         api_token = hv_token,
         hv_sites_arg = hv_sites,
         dump_dir = staging_directory
       )
     }
)
# 
# beep to let you know when it's done
beepr::beep(1)
```

## Pulling in any downloaded log data

Some logs (Vulink Log if sonde is AT600/800) cannot be backed up to HydroVu.
If you have any of these files, we will read them in here so that we are not missing any potential data.
Once this has been run and data checked, comment this section out to avoid re-running it.

```{r}
log_files <- list.files(log_directory, pattern = "\\vulink.htm$", full.names = TRUE)%>%
  #only keep files with our sites in them
  keep(~any(str_detect(tolower(.x), sites)))

log_files_simp <- basename(log_files)%>%
  str_replace_all(".htm", "")%>%
  str_squish()
# 
log_data <- map(log_files, ~rvest::read_html(.x)%>%
                            parse_insitu_html_log())

#check for "unnamed" sites in log_data
names(log_data) <- log_files_simp

cleaned_log_data <- log_data%>%
imap(~ mutate(.x, log_name = .y))%>%
  rbindlist()%>%
  #fixing a few specific location names before passing to fix_site_names
  mutate( site = tolower(site),
          id = as.double(sensor_sn),
         timestamp = with_tz(DT, tzone = "UTC"),
         site = case_when(site == "archery range"~ "archery",
                          site == "legacy bridge" ~ "salyer",
                         site == "boxelder creek" ~ "boxcreek",
                         site == "boxeldercreek" ~ "boxcreek",
                         site == "spring creek" ~ "springcreek",
                         site == "unnamed" ~ "salyer", # single salyer log with unnamed selected
                         TRUE ~ site),
         name = paste0("Vulink Log - ", site),
         #Correcting for parse log where temp is saved with degrees C rather than C
         units = case_when(parameter == "Temperature" ~ "C",
                           TRUE ~ unit)
         )%>%
  #make sure no erroneous dates (pre 2020 get through)
  filter(timestamp >= ymd_hms("2020-01-01 00:00:00", tz = "UTC"))%>%
  fix_site_names()%>%
  select(site, id, name, timestamp, parameter, value, units) %>%
  distinct(.keep_all = TRUE) %>%
  split(f = list(.$site), sep = "-") %>%
  keep(~nrow(.) > 0)

#write out the log data to the staging directory so it can be processed with the HV data
iwalk(cleaned_log_data, ~write_parquet(.x, here(staging_directory, paste0(.y, "_log_data.parquet"))))

# beep to let you know when it's done
beepr::beep(1)

```

### South Fork data

This data was logged on the WET API servers but did not log correctly so we will read it in seperately

```{r}
sf_data <- read_rds("data/raw/sensor/log_download/2023/SF_data_archive2023.RDS")%>%
  mutate(timestamp = with_tz(DT_round, "UTC"), 
         site = "sfm", 
         id = 112233, 
         name = "SFM WET API", 
         parameter = case_when(Measurement == "Specific_Conductivity" ~ "Specific Conductivity",
                               Measurement == "Chl-a" ~ "Chl-a Fluorescence",
                               Measurement == "FDOM" ~ "FDOM Fluorescence",
         TRUE ~ Measurement),
         value = Value, 
         units = case_when(
           parameter == "DO" ~ "mg/L", 
           parameter %in% c("Chl-a Fluorescence", "FDOM Fluorescence") ~ "RFU", 
           parameter %in% c("ORP") ~ "V", 
           parameter %in% c("pH") ~ "pH", 
           parameter %in% c("Specific Conductivity") ~ "µS/cm",
           parameter %in% c("Depth") ~ "m",
           parameter %in% c("Turbidity") ~ "NTU",
           parameter %in% c("Temperature") ~ "C", 
           TRUE ~ NA
         ))%>%
  fix_site_names()%>%
  select(site, id, name, timestamp, parameter, value, units) %>%
  distinct(.keep_all = TRUE) 

 write_parquet(sf_data, here(staging_directory, "sfm_log_data.parquet"))

```

# Raw Data Processing

This is where we will read in all the pulled data from HydroVu and our parsed logs to then be tidied and checked.

## Load in and tidy the raw files

Here we use `munge_api_data()` and `tidy_api_data()` functions from `ross.wq.tools` to read in and tidy the raw data files.
Our output will include lists seperated by site-parameter combinations for easier processing downstream.
Each list has DT_round (in UTC), site, parameter, units, mean, n_obs, spread, DT_join and flag columns and is summarized to a 15 minute interval

```{r}
#Since we no longer have any virridy sites in the network, we can just use a normal munge!
hv_data <- munge_api_data(api_dir = staging_directory,synapse_env = F)%>%
  #fix the virridy names to be their own sonde
  mutate(site = if_else(grepl("virridy", name, ignore.case = T), paste0(site, "_virridy"), site))%>%
  fix_site_names()%>%
  # Only keep parameters we are interested in
  filter(parameter %in% parameters)%>%
  # split into site-parameter combinations for parallel processing
  split(f = list(.$site, .$parameter), sep = "-") %>%
   keep(~nrow(.) > 0)


tidy_data <- hv_data %>%
  future_map(~tidy_api_data(api_data = .),.progress = TRUE, .options = set_furr_options) %>%  # the summarize interval default is 15 minutes
  keep(~!is.null(.))

```

## Double check that pulls were successful

Double check that the general data ranges appear to be correct and no sites or parameters were lost along the way.

```{r}
tidy_data%>%
  bind_rows()%>%
  summarise(date_range = paste(range(DT_round, na.rm = T), collapse = "---"),
            n_obs = n(),
            #get me a list of all the parameters in the data
            parameters = paste(unique(parameter), collapse = ", "), .by = "site")
```

We are giving it a quick look over to make sure there aren't any large data gaps (\>1 day) in the data pulled.
We are also adding in our field notes to the plots to make sure the data we have lines up with the periods we expect the sondes to be deployed and recording data.

```{r}
data_check_plots <- tidy_data%>%
  bind_rows()%>%
  #Summarizing to an hour to speed up plotting
  mutate(DT_hourly = floor_date(DT_round, unit = "1 hour"))%>%
  group_by(site, parameter, DT_hourly)%>%
  summarize(hourly_med = median(mean,na.rm = T ))%>%
  split(f = list(.$site))%>%
  map(., function(df) {
    
    field_notes <- all_field_notes %>%filter(site == unique(df$site) & year(DT_round) == "2023")%>%
      select(DT_round,sonde_employed)%>%
      mutate(sonde_changed = case_when(sonde_employed == 0 ~ "Deployed",
                                       sonde_employed == 1 ~ "Pulled",
                                       TRUE ~ "Active"))
    malfunction_notes <- sensor_malfunction_notes %>%filter(site == unique(df$site) & year(start_DT) == "2023")%>%
      select(start_DT, end_DT, parameter)%>%
      mutate(parameter = ifelse(parameter == "FDOM",  "FDOM Fluorescence", parameter))
    
    ggplot(df) +
      geom_point(aes(x = DT_hourly, y = hourly_med)) +
      geom_vline(data = field_notes, aes(xintercept = DT_round, color = sonde_changed))+
      geom_rect(data = malfunction_notes, aes(xmin = start_DT, xmax = end_DT, ymin = -Inf, ymax = Inf), fill = "gray", alpha = 0.8)+
      facet_wrap(~parameter, scales = "free_y")+
      labs(title = unique(df$site),
           x = "Date",
           y = "Hourly Median Value", 
           color = "Sonde Status")
  })
#Plot
walk(.x = data_check_plots, .f = print)

```

If there are discrepancies in the data, double check the raw files downloaded from HydroVu and the log files to make sure everything looks correct.
Occasionally there are issues with the HydroVu download missing data and that can be rerun for an individual site.
If there continue to be large gaps in data (\>3 days) during deployment periods, check field notes or troubleshooting may be needed

# Apply Back-calibration as needed

This uses the Shiny App (`calibration_correction_tool`) developed by Juan.
We are essentially deciding when a sensor needs to be back-calibrated based on field notes and applying the back-calibration to the raw data before flagging.
The majority of calibrations are fine as is but we want to "skip" over the bad ones.

## Read in calibration data

```{r}
#This creates calibration_data list object needed for the back-calibration functions
#NOTE: This will take a good while to run if update = T
load_calibration_data(
  cal_data_file_path = here::here("data", "collated", "sensor", "cal_reports", "munged_calibration_data.RDS"),
    update = FALSE, # set to TRUE when running for the first time
  field_cal_dir = here::here("data", "raw", "sensor", "calibration_reports")
) 


```

### Organize sensor data to match formatting for `cal_join_sensor_calibration_data()`

Data should be in a nested list structure: `sensor_data$years$site-parameter`

```{r}
sensor_data <- tidy_data %>%
  bind_rows()%>%
  #filter out virridy data bc they were not calibrated for this year
  filter(!grepl("virridy", site, ignore.case = T))%>%
  split(f = year(.$DT_round)) %>%
  map(~ {
    split_data <- split(.x, f = list(.x$site, .x$parameter), sep = "-")
    discard(split_data, ~is.null(.) || nrow(.) == 0)
  })

```

### Join to calibration data and apply back-calibration

This workflow follows the workflow outlined in the `ross.wq.tools` vignette on back calibration.
First, we join the calibration and sensor data so that each time instance has the slope and intercept info needed for back-calibration.
Next, we prepare the calibration windows to split the data into calibration windows.
Lastly we loop through each calibration window to apply the back-calibration using `cal_back_calibrate()`.

```{r}

sensor_calibration_data <- cal_join_sensor_calibration_data(
  sensor_data_list = sensor_data,
  calibration_data_list = calibration_data
 )

prepped_snsr_cal_data <- cal_prepare_calibration_windows(
  sensor_calibration_data_list = sensor_calibration_data
 )

calibrated_data <- prepped_snsr_cal_data%>%
  map(function(year){
    year%>%
      map(function(site_param){
        site_param %>%
          map_dfr(function(chunk){
            cal_back_calibrate(chunk)
          })
      })
  })

beepr::beep(1)

```

## Save to back calibrated folder

This will become the "original" file in the shiny app for manual verification

```{r}

 back_cal_dir <- here("data", "raw", "sensor", "manual_data_verification", "2023_cycle","hydro_vu_pull","back_calibration")

if(!dir_exists(back_cal_dir)){
  dir_create(back_cal_dir)
}

 write_rds(calibrated_data, here(back_cal_dir, "calibrated_sensor_data_2023.rds"))
 write_rds(new_tracking, here(back_cal_dir, "calibrated_sensor_field_data_tracking.rds"))
```

Take \^ file and run through the shiny app to verify back-calibration and make any manual adjustments as needed.

# Run back calibration shiny app

```{r}

# Load required packages for ShinyApp
 invisible(lapply(c("shiny","bslib","DT","tidyverse","lubridate","here","ggpubr","gridExtra","plotly",
                   "keys","patchwork","digest","fs","shinyFiles","shinyWidgets","glue","anytime","arrow"),
                 package_loader)
           )
```

Run this code chunk to launch the calibration correction tool.
The app will open in a new window or tab in your default web browser.

```{r, echo=FALSE}
shiny::shinyAppDir(
  appDir = here("calibration_correction_tool"),
  options = list(width = "100%", height = "100%")
 )
```

## Read in user corrected back-calibrated data

```{r}
 calibrated_data_final <- read_rds(here("data", "raw", "sensor", "manual_data_verification", "2023_cycle","hydro_vu_pull","back_calibration", "calibrated_sensor_field_data_finalized.rds"))[["2023"]] # Only extract the 2023 data

```

## Quick look at the back-calibrated data

All of this should already have been double checked by the user but we just want to make sure everything looks good before proceeding to flagging.
This plot will show the raw vs back-calibrated data ("final") with vertical lines indicating when the calibration events occurred.

```{r}

plots <- map(names(calibrated_data_final),
             function(site_param) {
               plot_data <- calibrated_data_final[[site_param]]

               vline_df <- plot_data %>%
                 group_by(sensor_date) %>%
                 slice_min(DT_round, n = 1) %>%
                 arrange(DT_round)%>%
                 ungroup()%>%
                 select(DT_round, file_date)%>%
                 mutate(file_date_mst = with_tz(file_date, tzone = "MST"),
                        file_date_mst_utc = force_tz(file_date_mst, tzone = "UTC"))

               param <- unique(plot_data$parameter)
               site <- unique(plot_data$site)

               ggplotly(ggplot(plot_data, aes(DT_round))+
                          geom_line(aes(y = mean, color = "raw"))+
                          geom_line(aes(y = mean_cal, color = "final"))+
                          geom_vline(xintercept = vline_df$file_date_mst_utc, linetype = "dashed")+
                          scale_color_manual(name = "Data Type", values = c("raw" = "blue", "final" = "green4"))+
                          labs(title = paste0(site, " - ", param),
                               x = "Date",
                               y = param))
             })%>%
   set_names(names(calibrated_data_final))

#Typically easiest to go parameter-wise through the plots
 plots[grepl("pH", names(plots))]

```

## Remove columns as needed

Here we are removing columns that are not needed downstream to reduce memory usage.
We are also merging back in the units, n_obs, spread and flag columns from the original tidy data.
Temperature and Depth are not passed into the back-calibration process so we will grab those (and any other site-parameters that were not calibrated) from the original tidy data as well.
We will also add the column "back_cal_performed" to indicate whether a back-calibration was actually applied to the data or not.

```{r}
# 
cal_corrected_data <- calibrated_data_final%>%
    map(., function(df){

    site <- unique(df$site)
    parameter <- unique(df$parameter)

    site_param = paste0(site, "-", parameter)

    add_cols <- tidy_data[[site_param]]%>%
      select( DT_join, units, n_obs, spread, flag)
  

    if("updated_slope_from" %in% names(df)){
      df <- df%>%
        mutate(back_cal_performed = if_else(slope == updated_slope_to & slope == updated_slope_from, F, T))
    } else {
      df <- df%>%
        mutate(back_cal_performed = F)
    }

    df <- df%>%
      select(
        DT_round,
        site,
        parameter,
        mean_pre_cal = mean,
        mean = mean_cal,
        back_cal_performed
      )%>%
      #DT_join is dropped in the calibration process so we need to add it back in to join with the tidy_data columns
      #This is still in UTC so we can join with the tidy_data columns
      mutate(DT_join = as.character(DT_round))%>%
      left_join(add_data, by = c("DT_join"))
    #Some data is missed if there is not sufficient calibration files to create windows across all data
    #Here we can grab the data that was not back-calibrated and add it back in with the back_cal_performed column set to FALSE
    non_caled_data <- tidy_data[[site_param]]%>%
      filter(!(DT_join %in% df$DT_join))%>%
       mutate(back_cal_performed = FALSE, 
              mean_pre_cal = mean)
   df%>%
      bind_rows(non_caled_data)
# 
# # We will also need to grab the non-calibrated data (temperature and depth, virridy sites and other sites without calibration info) and add it back in with the back_cal_performed column set to FALSE

non_calibrated_names <- setdiff(names(tidy_data), names(calibrated_data_final))

non_cal_corrected_data <- tidy_data[non_calibrated_names]%>%
  map(., function(df){
    df%>%
      mutate(back_cal_performed = FALSE,
             mean_pre_cal = mean)%>%
      select(,
             DT_round,
             site,
             parameter,
             mean_pre_cal,
             mean,
             back_cal_performed,
             DT_join,
             units,
             n_obs,
             spread,
             flag
      )
  })

# 
# # Bind all data together
  tidy_cal_data <- c(cal_corrected_data, non_cal_corrected_data)

```

# Data Flagging

## Combine with field notes

Here we read in our field notes and then join them to the calibrated data using `add_field_notes()`.
We then generate summary statistics (slope, lag, etc) using `generate_summary_statistics()`.

```{r}

# Add the field note data to all of the data
all_field_notes <- all_field_notes %>%
  fix_site_names()
#Creating duplicate notes for virridy sites
virridy_duplicate_notes <- all_field_notes%>%
  filter(site %in% c("cottonwood", "riverbend", "archery") & DT_round >= "2023-09-01")%>%
  mutate(site = case_when(site == "cottonwood" ~ "cottonwood_virridy",
                          site == "riverbend" ~ "riverbend_virridy",
                          site == "archery" ~ "archery_virridy",
                          TRUE ~ site))

all_field_notes <- all_field_notes%>%
  bind_rows(virridy_duplicate_notes)

combined_data <- tidy_cal_data %>% 
  future_map(~add_field_notes(df = ., notes = all_field_notes), .progress = TRUE)

# Add summary statistics
summarized_data <- combined_data %>%
  map(~generate_summary_statistics(.))

```

### Final Double Check before flagging

Double check we have all the right site-parameters and ranges.

```{r}
summarized_data%>%
  bind_rows()%>%
  summarise(date_range = paste(range(DT_round, na.rm = T), collapse = "---"),
            n_obs = n(),
            #get me a list of all the parameters in the data
            parameters = paste(unique(parameter), collapse = ", "), .by = "site")
```

Double check we don't have any big missing gaps of data

```{r}
data_check_plots <- summarized_data%>%
  bind_rows()%>%
  #Summarizing to an hour to speed up plotting
  mutate(DT_hourly = floor_date(DT_round, unit = "1 hour"))%>%
  group_by(site, parameter, DT_hourly, )%>%
  summarize(hourly_med = median(mean,na.rm = T ), 
            hourly_med_pre_cal = median(mean_pre_cal,na.rm = T ))%>%
  split(f = list(.$site))%>%
  map(., function(df) {
    
    field_notes <- all_field_notes %>%filter(site == unique(df$site) & year(DT_round) == "2023")%>%
      select(DT_round,sonde_employed)%>%
      mutate(sonde_changed = case_when(sonde_employed == 0 ~ "Deployed",
                                       sonde_employed == 1 ~ "Pulled",
                                       TRUE ~ "Active"))
    malfunction_notes <- sensor_malfunction_notes %>%
      filter(site == unique(df$site) & year(start_DT) == "2023")%>%
      select(start_DT, end_DT, parameter)%>%
      mutate(parameter = ifelse(parameter == "FDOM",  "FDOM Fluorescence", parameter))%>%
      filter(parameter %in% unique(df$parameter))
    
ggplot(df) +
  geom_point(aes(x = DT_hourly, y = hourly_med, color = hourly_med == hourly_med_pre_cal)) +
  geom_vline(data = field_notes, aes(xintercept = DT_round, color = sonde_changed)) +
  geom_rect(data = malfunction_notes, aes(xmin = start_DT, xmax = end_DT, ymin = -Inf, ymax = Inf), fill = "gray", alpha = 0.8) +
  facet_wrap(~parameter, scales = "free_y") +
  labs(title = unique(df$site),
       x = "Date (UTC)",
       y = "Hourly Median Value",
       color = "Sonde Status / Back-calibrated?")
  })
#Plot
walk(.x = data_check_plots, .f = print)

```

## Single Sensor Flags

This is the first round of flagging that is done on a per-site-parameter basis.
Here we are applying functions to flag field visits, missing data, DO noise, repeating values, depth shifts, optical sensor drift, spec range violations and seasonal range violations.
This can take \~5-15 min depending on data volume.

```{r}
# Chunk the data for furrr
summarized_data_chunks <- split(1:length(summarized_data),
                                ceiling(seq_along(1:length(summarized_data))/10))
# Flag data...
# Single parameter flags
single_sensor_flags <- list()
for (chunk_idx in seq_along(summarized_data_chunks)) {
  message("\n=== Processing chunk ", chunk_idx, " of ", length(summarized_data_chunks), " ===")
  
  # Get the indices for this chunk
  indices <- summarized_data_chunks[[chunk_idx]]
  chunk_data <- summarized_data[indices]
  
  # Process the chunk in parallel
  chunk_results <- chunk_data %>%
    future_map(
      function(data) {
        flagged_data <- data %>%
          data.table(.) %>%
          # flag field visits
          add_field_flag(df = .) %>%
          # flag missing data
          add_na_flag(df = .) %>%
          # flag DO noise
          find_do_noise(df = .) %>%
          # flag repeating values
          add_repeat_flag(df = .) %>%
          # find times when sonde was moved up/down in housing
          add_depth_shift_flag(df = ., level_shift_table =  all_field_notes, post2024 = TRUE) %>%
          # find instances of sensor drift (FDOM, Chl-a, Turbidity only)
          add_drift_flag(df = .)
        
        if (unique(data$parameter) %in% names(sensor_thresholds)) {
          # flag instances outside the spec range
          flagged_data <- flagged_data %>%
            data.table(.) %>%
            add_spec_flag(df = ., spec_table = sensor_thresholds)
        }
        
        if (unique(data$parameter) %in% unique(season_thresholds$parameter)) {
          # flag instances outside the spec range
          flagged_data <- flagged_data %>%
            data.table(.) %>%
            add_seasonal_flag(df = ., threshold_table = season_thresholds)
        }
        
        flagged_data <- flagged_data %>%
          data.table(.)
        
        return(flagged_data)
      },
      .progress = TRUE, .options = set_furr_options
    )
  
  # Add chunk to list
  single_sensor_flags <- c(single_sensor_flags, chunk_results)
  
  if (chunk_idx < length(summarized_data_chunks)) {
    message("Taking a short break before next chunk...")
    gc()
    Sys.sleep(0.1)
  }
}

# beep to let you know when it's done
beepr::beep(1)

```

## Intrasensor flags

Now we will apply flags that require multiple parameters from the same sensor to be present.
This includes frozen water flagging (temp \< 0C), unsubmerged flagging (depth \< 0 or SC \< 10) and sonde burial flagging (long-term DO noise).
We also do an intersensor check to remove slope violation flags that occur concurrently with temp or depth violations to reduce overflagging.
This can take \~5-15 min depending on data volume.

```{r}

intrasensor_flags <- single_sensor_flags %>%
  rbindlist(fill = TRUE) %>%
  split(by = "site")

# Chunk the data for furrr
intrasensor_data_chunks <- split(1:length(intrasensor_flags),
                                 ceiling(seq_along(1:length(intrasensor_flags))/3))

intrasensor_flags_list <- list()
for (chunk_idx in seq_along(intrasensor_data_chunks)) {
  message("\n=== Processing chunk ", chunk_idx, " of ", length(intrasensor_data_chunks), " ===")
  
  # Get the indices for this chunk
  indices <- intrasensor_data_chunks[[chunk_idx]]
  chunk_data <- intrasensor_flags[indices]
  # Process the chunk in parallel
  chunk_results <- chunk_data %>%
    future_map(
      function(data) {
        # A chunk is a site df
        flagged_data <- data %>%
          data.table() %>%
          # flag times when water was below freezing
          add_frozen_flag(.) %>%
          # overflagging correction. remove slope violation flag if it occurs concurrently
          # with temp or depth
          intersensor_check(.) %>%
          # add sonde burial. If DO is noise is long-term, likely burial:
          add_burial_flag(.) %>%
          # flag times when sonde was unsubmerged
          add_unsubmerged_flag(.)
        
        return(flagged_data)
      }, .progress = TRUE, .options = set_furr_options
    ) %>%
    rbindlist(fill = TRUE) %>%
    # lil' cleanup of flag column contents
    dplyr::mutate(flag = ifelse(flag == "", NA, flag)) %>%
    distinct(site, parameter, DT_round, mean, .keep_all = TRUE) %>%
    # transform back to site-parameter dfs
    split(f = list(.$site, .$parameter), sep = "-") %>%
    purrr::discard(~ nrow(.) == 0) %>%
    # Add in KNOWN instances of sensor malfunction
    map(~add_malfunction_flag(df = ., malfunction_records = sensor_malfunction_notes))
  
  # Add chunk to list
  intrasensor_flags_list <- c(intrasensor_flags_list, chunk_results)
  
  if (chunk_idx < length(intrasensor_data_chunks)) {
    message("Taking a short break before next chunk...")
    gc()
    Sys.sleep(0.1)
  }
}

# beep to let you know when it's done
beepr::beep(1)

```

### Let's temporarily save this data so we can remove everything else from memory

```{r}
iwalk(intrasensor_flags_list, ~write_parquet(.x, here(flagged_temp_dir, paste0(.y, ".parquet"))))

#remove everything other than intrasensor flags from memory
rm(list=setdiff(ls(), c("base_folder", "intrasensor_flags_list", "flagged_directory")))
#clean up memory
gc()

```

## Applying network check

The network check function `ross.wq.tools::network_check()` compares data across sites to identify when data is being overflagged and remove flags as needed.
This step requires knowledge of the site order, which is read in from a YAML file and should be updated as sites change.

```{r}
#Read in site order for year/pull
site_order_2023 <- load_site_order( here("data","derived","auto_qaqc_files", "site_order", "PWQN_site_order_2023.yml"))

#Applying Network check (non parrallel so only map used)
final_flags <- intrasensor_flags_list %>%
  map(~network_check(df = ., intrasensor_flags_arg = intrasensor_flags_list, site_order_arg = site_order_2023)) %>%
  rbindlist(fill = TRUE) %>%
  tidy_flag_column() %>%
  split(f = list(.$site, .$parameter), sep = "-") %>%
  map(~add_suspect_flag(.)) %>%
  rbindlist(fill = TRUE)

v_final_flags <- final_flags%>%
  mutate(auto_flag = ifelse(is.na(auto_flag), NA,
                            ifelse(auto_flag == "suspect data" & is.na(lag(auto_flag, 1)) & is.na(lead(auto_flag, 1)), NA, auto_flag))) %>%
  select(c("DT_round", "DT_join", "site", "parameter",
           "mean_pre_cal","mean", "units", "n_obs", "spread", "auto_flag", "mal_flag", "sonde_moved","sonde_employed", "season", "last_site_visit", "depth_change","back_cal_performed")) %>%
  mutate(auto_flag = ifelse(is.na(auto_flag), NA, ifelse(auto_flag == "", NA, auto_flag))) %>%
  split(f = list(.$site, .$parameter), sep = "-") %>%
  keep(~nrow(.) > 0)

# beep to let you know when it's done
beepr::beep(1)

```

## Check final output

Here we are making sure that all sites data were flagged and giving a quick look at the flagged data to make sure everything looks correct.
With newer sites/parameters, we expect there to be over-flagging as the seasonal ranges/slopes are going to be very small.

```{r}
v_final_flags%>%
  bind_rows()%>%
  split(f = list(.$site))%>%
  map(., function(df) {
    ggplot(df, aes(x = DT_round, y = mean, color = auto_flag)) +
      geom_point() +
      facet_wrap(~parameter, scales = "free_y")+
      labs(title = unique(df$site),
           x = "Date",
           y = "Mean Value")
  }) %>%
  walk(plot)
```

# Save Final Flagged Dataset

We will save it into three places:

- Individually flagged site-parameter parquet files for posterity and comparison with manual verification to determine effectiveness of auto-flagging

- A combined raw data parquet file for use in the `manual_verification_tool` shiny app for posterity

- Individual site-parameter files to populate the pre and all data directories for the shiny app

## Save to flagged directory

```{r}

# Save the data individually.
iwalk(v_final_flags, ~write_parquet(.x, here(flagged_directory,  paste0(.y, ".parquet"))))

```

## Save for manual verification shiny app

```{r}
base_dir = here("data", "raw", "sensor", "manual_data_verification", "2023_cycle", "in_progress")
  all_path = here(base_dir, "all_data_directory")
  pre_verification_path = here(base_dir, "pre_verification_directory")
  intermediary_path = here(base_dir, "intermediary_directory")
  verified_path = here(base_dir, "verified_directory")
  raw_data_path = here(base_dir, "raw_data")
  meta_folder_path = here(base_dir, "meta")
  
  
  #check/ create all necessary folders
  if(!dir.exists(all_path)){
    dir.create(all_path)
  }
  #check to see if data folder has pre_verification_directory
  if(!dir.exists(pre_verification_path)){
    dir.create(pre_verification_path)
  }
  #check to see if data folder has intermediary_directory
  if(!dir.exists(intermediary_path)){
    dir.create(intermediary_path)
  }
  
  #check to see if data folder has verified_directory
  if(!dir.exists(verified_path)){
    dir.create(verified_path)
  }
  #create raw data folder
  if(!dir.exists(raw_data_path)){
    dir.create(raw_data_path)
    write_parquet(v_final_flags%>%bind_rows(), here(raw_data_path, "raw_data_2023.parquet"))
  }
  if(!dir.exists(meta_folder_path)){
    dir.create(meta_folder_path)
    #save the site order file and seasonal threshold file to the meta folder for use in the app
    file.copy(from = here("data", "derived", "auto_qaqc_files", "site_order", "PWQN_site_order_2023.yml"), to = meta_folder_path)
    file.copy(from = here("data","derived","auto_qaqc_files", "thresholds","seasonal_thresholds20260304-T175010Z.csv"), to = meta_folder_path)
    file.rename(from = here(meta_folder_path, "seasonal_thresholds20260304-T175010Z.csv"), to = here(meta_folder_path, "seasonal_thresholds.csv"))
    #copy over the available flags file to the meta folder for use in the app
    file.copy(from = here("data","raw","sensor", "manual_data_verification","2025_cycle", "in_progress", "meta",   "available_flags.csv"), to = meta_folder_path)
    #copy over the parameter auto selections
    file.copy(from = here("data","raw","sensor", "manual_data_verification","2025_cycle", "in_progress", "meta",   "parameter_autoselections.csv"), to = meta_folder_path)
  }
  
  all_data_parsed <- v_final_flags%>%
    bind_rows()%>%
    mutate(DT_round = with_tz(DT_round, tzone = "MST"))%>%
       mutate(year = year(DT_round))%>%
    #double check we are only saving one year
    filter(year == "2023")
  
  #Taking all_data_parsed and formatting to save into all and pre directories
  site_params <- expand_grid(site = unique(all_data_parsed$site), parameter = unique(all_data_parsed$parameter))
  #create a list were each is named after the site and parameter combination
  flagged_list <- pmap(site_params, ~ all_data_parsed %>% filter(site == ..1, parameter == ..2))
  rm(all_data_parsed) #save some memory here
  names(flagged_list) <- paste(site_params$site, site_params$parameter, sep = "-")
  #remove empty lists
  flagged_list <- keep(flagged_list, ~nrow(.x) > 0)
  #add additional columns needed
  pre_processed_data <- map(.x = flagged_list, \(x)
                            x %>%
                              mutate(
                                mean_verified = mean, # if verification status pass or flag, this is mean. set default to mean so that it can be used in the app
                                is_verified = FALSE, # whether or not the data has been verified
                                verification_status = NA, # can only be pass/fail/skip
                                year = year(DT_round),
                                week = week(DT_round),
                                weekday = lubridate::wday(DT_round, week_start = 7),
                                y_w = paste(year, "-", week),
                                day = yday(DT_round),
                                y_d = paste(year, "-", day),
                                #based on Colorado Hydrology, should be removed for DS version
                                season = case_when(
                                  month(DT_round) %in% c(12, 1, 2, 3, 4) ~ "winter_baseflow",
                                  month(DT_round) %in% c(5, 6) ~ "snowmelt",
                                  month(DT_round) %in% c(7, 8, 9) ~ "monsoon",
                                  month(DT_round) %in% c(10, 11) ~ "fall_baseflow"
                                ),
                                #auto_flag = flag,
                                #merge auto_flag and mal_flag by ; into flag. don't include NAs
                                flag = case_when(
                                  #both are flagged then combine by ;
                                  !is.na(auto_flag)&!is.na(mal_flag) ~ paste(auto_flag, mal_flag, sep = ";"),
                                  #only auto_flag is flagged, only use auto_flag
                                  !is.na(auto_flag)&is.na(mal_flag) ~ auto_flag,
                                  #only mal_flag is flagged, only use mal_flag
                                  !is.na(mal_flag)&is.na(auto_flag) ~ mal_flag
                                ),
                                user_flag = flag, # Keep Auto flags in flag column, user added flags/alterations live in user flag
                                brush_omit = FALSE, # User Brush Omit instances, default to FALSE
                                user = NA, #User initials
                                final_status = NA, # Final status of the data point after weekly decision
                                week_decision = NA,  #store weekly decision
                                is_finalized = FALSE) #Only TRUE if user submits final decision button
  )
  rm(flagged_list) #save some memory here

```

### Final double check for TZs

We should expect to see a diurnal pattern in the temperature data, with peaks around \~ 15:00 (3pm) and troughs around \~ 5:00/7:00 (5-7am).
If we don't see this pattern, it could be an indication that the timestamps are not correct or that there is some other issue with the data.

```{r}
#Double check TZs, grab all temperature data and plot monthly averages by hour for each site
  pre_processed_data%>%
    keep(~ unique(.x$parameter) %in% c("Temperature"))%>%
    bind_rows()%>%
    mutate(hour = hour(DT_round))%>%
    group_by(site, parameter, month = month(DT_round), hour)%>%
    summarize(mean = mean(mean, na.rm = TRUE))%>%
    ggplot(aes(x = hour, y = mean, color = site))+
    geom_line()+
    facet_wrap(~month)
```

### Save files

```{r}
  
  iwalk(pre_processed_data, \(x, idx) {
    timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
    data_hash <- digest::digest(x)
    new_filename <- glue::glue("{idx}_{timestamp}_{data_hash}.parquet")
    write_parquet(x, here(all_path, new_filename))  # Ensure you use new_filename, not idx
  })
  
  R.utils::copyDirectory(all_path, pre_verification_path)

```

# Pass data through Manual Verification Tool

This tool is located in the folder `manual_verification_tool`.
Here we are manually verifying the data, making any necessary corrections and then saving the output in the "verified_directory" folder for this year.
The key columns added are:

- final_status:

  -  "PASS" if the user has verified the data as good

  - "FLAGGED" if the user has flagged the data for suspect but still useable data

  -  "OMIT if the user has concluded that the data should not be used in analysis due to sensor error

- user_flag: more specific flag added by the user to indicate why they flagged/omitted the data

  - (e.g. "suspect data", "sonde burial", "frozen water", "unsubmerged", "depth shift", "drift", "field visit", "other malfunction" etc).

  -  This is for posterity and to help inform auto-flagging performance.

- mean_analysis: this is the values where the "OMIT" final status was not applied.

```{r shiny-dependencies}
# Load required packages for ShinyApp
invisible(lapply(c("shiny","bslib","DT","tidyverse","lubridate","here","ggpubr","gridExtra","plotly",
                   "keys","patchwork","digest","fs","shinyFiles","shinyWidgets","glue","anytime","arrow"),
                 package_loader)
          )
```

Run this codechunk to launch the manual verification tool.
The app will open in a new window or tab in your default web browser.

```{r, echo=FALSE}
shiny::shinyAppDir(
  appDir = here("manual_verification_tool"),
  options = list(width = "100%", height = "100%")
)
```

# Pass data through Drift Correction Tool

This tool is located in the script `drift_correction_tool`.
Here we are applying the drift corrections identified in the manual verification tool to the data and then saving the output in the "post_verification" folder.

The key columns added are:

- mean_drift_trans: new data column with drift corrected values

- correction_type: what kind of drift correction was applied (exponential, fitted linear, etc)

  - This can also be "non_resolved" if the drift could not be appropriately solved via the current methods.
    These data are then moved from final_status = "FLAGGED" to final_status = "OMIT"

- pre_post_source: A, B, C depending on where the correction values came from

  - Outlined in `data_organization/drift_correction_2024.rmd`

# Prep for HydroShare Upload

## Depth Corrections

If we have a site visit where depth was adjusted up or down, we want to try to correct that.
Ideally we do this around the "Apply Back Calibration" step but that was not done for this year.

### Function

```{r}

adjust_depth_window <- function(data, 
                                site_oi,
                                win_start, 
                                win_end, 
                                dt_before, 
                                dt_start, 
                                site_col = "site",
                                tz = "MST") {
  #Double check that the tz matches the data$DT_round tz
  if (tz(data$DT_round) != tz) {
    stop(paste("Timezone mismatch: data$DT_round is in", tz(data$DT_round), "but tz argument is", tz))
  }
  #Double check that this site exists in the data
  if (!site_oi %in% unique(data[[site_col]])) {
    stop(paste("Site", site_oi, "not found in data. Available sites:", paste(unique(data[[site_col]]), collapse = ", ")))
  }  
  
  w_start   <- ymd_hm(win_start, tz = tz)
  w_end     <- ymd_hm(win_end, tz = tz)
  t_before  <- ymd_hm(dt_before, tz = tz)
  t_start   <- ymd_hm(dt_start, tz = tz)
  
  # Ensure target column exists; initialize with 'mean' if it's the first run
  if (!"adj_depth" %in% names(data)) {
    data <- data %>% 
      mutate(adj_depth = mean)%>%
      mutate(depth_corrected = FALSE)
  }
  
  # Extract values specifically for this site using current state of adj_depth
  site_mask <- data[[site_col]] == site_oi
  val_before <- data$adj_depth[site_mask & data$DT_round == t_before]
  val_start  <- data$adj_depth[site_mask & data$DT_round == t_start]
  
  # Safety Check: check if timestamps exist
  if (length(val_before) == 0 || length(val_start) == 0) {
    warning(paste("Target timestamps missing for site:", site, "- Skipping."))
    return(data)
  }
  
  offset <- val_before[1] - val_start[1]
  
  # Update ONLY the matching rows, leaving all other sites/rows untouched
  data %>%
    mutate(
      adj_depth = if_else(
        .data[[site_col]] == site_oi & DT_round >= w_start & DT_round <= w_end & final_status != "OMIT",
        adj_depth + offset,
        adj_depth
      ), 
      depth_corrected = if_else(
        .data[[site_col]] == site_oi & DT_round >= w_start & DT_round <= w_end,
        TRUE,
        depth_corrected
      ), 
      user_flag = if_else(
        .data[[site_col]] == site_oi & DT_round >= w_start & DT_round <= w_end,
        gsub("calibration error", "", user_flag),
        user_flag
      ), 
      user_flag = if_else(user_flag == "", NA_character_, user_flag), 
      #Update decision to "PASS"
      final_status = if_else(
        .data[[site_col]] == site_oi & DT_round >= w_start & DT_round <= w_end & final_status != "OMIT",
        "PASS",
        final_status
      )
    )
}

```

### Applying

```{r}

depth_files <- list.files(path = 'data/raw/sensor/manual_data_verification/2023_cycle/in_progress/verified_directory/', pattern = "*.parquet", full.names = TRUE)%>%
  grep(pattern = "depth", x = ., ignore.case = TRUE,value = T)

depth_data <- map_dfr(depth_files, read_parquet)%>%
  # Bellvue Depth Adjustment (June - July): Two Adjustments
  adjust_depth_window(
    site_oi      = "bellvue",
    win_start = "2023-06-08 12:30",
    win_end   = "2023-06-26 08:30",
    dt_before = "2023-06-08 11:45",
    dt_start  = "2023-06-08 13:00"
  ) %>%
  adjust_depth_window(
    site_oi      = "bellvue",
    win_start = "2023-06-26 11:45",
    win_end   = "2023-07-17 14:00",
    dt_before = "2023-06-26 08:30",
    dt_start  = "2023-06-26 11:45"
  )%>%
  #Riverbend Depth Adjustments
  adjust_depth_window(
    site_oi      = "riverbend",
    win_start = "2023-06-14 11:45",
    win_end   = "2023-06-26 08:00",
    dt_before = "2023-06-13 13:15",
    dt_start  = "2023-06-14 12:00"
  ) %>%
  adjust_depth_window(
    site_oi      = "riverbend",
    win_start = "2023-06-26 11:45",
    win_end   = "2023-06-27 8:15",
    dt_before = "2023-06-26 08:00",
    dt_start  = "2023-06-26 11:45"
  )%>%
   adjust_depth_window(
    site_oi      = "riverbend",
    win_start = "2023-06-27 9:15",
    win_end   = "2023-07-18 15:15",
    dt_before = "2023-06-27 08:15",
    dt_start  = "2023-06-27 09:15"
  )%>%
    adjust_depth_window(
    site_oi      = "riverbend",
    win_start = "2023-09-12 13:00",
    win_end   = "2023-09-17 9:30",
    dt_before = "2023-09-12 12:15",
    dt_start  = "2023-09-12 13:00"
  )%>%
   adjust_depth_window(
    site_oi      = "riverbend",
    win_start = "2023-10-31 15:15",
    win_end   = "2023-11-02 15:30",
    dt_before = "2023-10-31 13:45",
    dt_start  = "2023-10-31 15:15"
  )%>%
  #elc Depth Adjustments
  adjust_depth_window(
    site_oi      = "elc",
    win_start = "2023-06-28 9:15",
    win_end   = "2023-06-29 08:45",
    dt_before = "2023-06-28 08:00",
    dt_start  = "2023-06-28 9:15"
  ) %>%
  adjust_depth_window(
    site_oi      = "elc",
    win_start = "2023-06-29 10:45",
    win_end   = "2023-07-17 08:30",
    dt_before = "2023-06-29 08:45",
    dt_start  = "2023-06-29 10:45"
  ) %>%
   adjust_depth_window(
    site_oi      = "elc",
    win_start = "2023-07-19 13:45",
    win_end   = "2023-08-10 09:45",
    dt_before = "2023-07-19 12:45",
    dt_start  = "2023-07-19 13:45"
  ) %>%
  #joei Depth Adjustments
   adjust_depth_window(
    site_oi      = "joei",
    win_start = "2023-10-05 09:00",
    win_end   = "2023-10-10 11:45",
    dt_before = "2023-10-05 8:45",
    dt_start  = "2023-10-05 09:00"
  ) %>%
  #Udall Depth Adjustments
   adjust_depth_window(
    site_oi      = "udall",
    win_start = "2023-10-20 13:30",
    win_end   = "2023-10-26 13:00",
    dt_before = "2023-10-20 13:00",
    dt_start  = "2023-10-20 13:30"
  ) %>%
    #Riverbluffs Depth Adjustments
   adjust_depth_window(
    site_oi      = "riverbluffs",
    win_start = "2023-05-08 17:45",
    win_end   = "2023-05-10 15:30",
    dt_before = "2023-05-08 16:30",
    dt_start  = "2023-05-08 17:45"
  ) %>%
   adjust_depth_window(
    site_oi      = "riverbluffs",
    win_start = "2023-05-16 12:45",
    win_end   = "2023-05-22 15:45",
    dt_before = "2023-05-16 11:45",
    dt_start  = "2023-05-16 12:45"
  ) %>%
  adjust_depth_window(
    site_oi      = "riverbluffs",
    win_start = "2023-05-22 16:30",
    win_end   = "2023-06-27 12:00",
    dt_before = "2023-05-22 15:45",
    dt_start  = "2023-05-22 16:30"
  ) %>%
  adjust_depth_window(
    site_oi      = "riverbluffs",
    win_start = "2023-06-27 13:45",
    win_end   = "2023-06-29 10:30",
    dt_before = "2023-06-27 12:00",
    dt_start  = "2023-06-27 13:45"
  ) %>%
  # Riverbend Virridy Adjustments
  adjust_depth_window(
    site_oi      = "riverbend_virridy",
    win_start = "2023-10-31 15:15",
    win_end   = "2023-11-02 15:30",
    dt_before = "2023-11-02 15:45",
    dt_start  = "2023-11-02 15:30"
  ) %>%
  #remove any instances where the final status is OMIT from the mean_analysis column so that they are not used in analysis
  mutate(mean_analysis = ifelse(final_status  %in% c("FLAGGED", "PASS"), adj_depth, NA))


final_depth_data<- depth_data%>%
  select(DT_round, DT_join, site, parameter,units, 
         value_raw = mean_pre_cal, #orginal raw data
         value_post_cal_corr = adj_depth, # calibration corrected depth (code above)
         value_post_ver = mean_analysis, # Omit data removed
         value_post_drift_corr = mean_analysis,   # same as omit data removed
         auto_qc_flag = auto_flag, mal_flag, 
         user_qc_flag = user_flag, qc_status = final_status)


# ggplotly(ggplot(final_depth_data%>%filter(site %in% c( "riverbend")), aes(x = DT_round, y = value_raw))+
#   geom_point()+
#   geom_point(aes(y = value_post_ver, color = qc_status),  alpha = 1)+
#   #geom_line(aes(y = value_post_cal_corr), color = "black", alpha = 0.5)+
#   
#   labs(title = "Depth Data with Adjustments",
#        x = "Date",
#        y = "Depth (m)",
#        color = "Depth Corrected")+
#   facet_wrap(~site, scales = "free_y"))

```

## Turbidity Offset Corrections

A few sites have turbidity values that are miscalibrated and primarily need a small offset to correct.
This was not corrected in calibration correction tool

### Function

```{r}
adjust_turbidity_window <- function(data, 
                                site_oi,
                                win_start, 
                                win_end, 
                                dt_before, 
                                dt_start, 
                                site_col = "site",
                                tz = "MST") {
  #Double check that the tz matches the data$DT_round tz
  if (tz(data$DT_round) != tz) {
    stop(paste("Timezone mismatch: data$DT_round is in", tz(data$DT_round), "but tz argument is", tz))
  }
  #Double check that this site exists in the data
  if (!site_oi %in% unique(data[[site_col]])) {
    stop(paste("Site", site_oi, "not found in data. Available sites:", paste(unique(data[[site_col]]), collapse = ", ")))
  }  
  
  w_start   <- ymd_hm(win_start, tz = tz)
  w_end     <- ymd_hm(win_end, tz = tz)
  t_before  <- ymd_hm(dt_before, tz = tz)
  t_start   <- ymd_hm(dt_start, tz = tz)
  
  # Ensure target column exists; initialize with 'mean' if it's the first run
  if (!"adj_turb" %in% names(data)) {
    data <- data %>% 
      mutate(adj_turb = mean_drift_trans)%>%
      mutate(turb_corrected = FALSE)
  }
  
  # Extract values specifically for this site using current state of adj_depth
  site_mask <- data[[site_col]] == site_oi
  val_before <- data$adj_turb[site_mask & data$DT_round == t_before]
  val_start  <- data$adj_turb[site_mask & data$DT_round == t_start]
  
  # Safety Check: check if timestamps exist
  if (length(val_before) == 0 || length(val_start) == 0) {
    warning(paste("Target timestamps missing for site:", site, "- Skipping."))
    return(data)
  }
  
  offset <- val_before[1] - val_start[1]
  
  # Update ONLY the matching rows, leaving all other sites/rows untouched
  data %>%
    mutate(
      adj_turb = if_else(
        .data[[site_col]] == site_oi & DT_round >= w_start & DT_round <= w_end,
        adj_turb + offset,
        adj_turb
      ), 
      turb_corrected = if_else(
        .data[[site_col]] == site_oi & DT_round >= w_start & DT_round <= w_end,
        TRUE,
        turb_corrected
      ), 
      user_flag = if_else(
        .data[[site_col]] == site_oi & DT_round >= w_start & DT_round <= w_end,
        gsub("calibration error", "", user_flag),
        user_flag
      ), 
      user_flag = if_else(user_flag == "", NA_character_, user_flag)
    )
}
```

### Applying

```{r}
turb_files <- list.files(path = 'data/raw/sensor/manual_data_verification/2023_cycle/post_verification/', pattern = "*.parquet", full.names = TRUE)%>%
  grep(pattern = "Turbidity", x = ., ignore.case = TRUE,value = T)

turb_data <- map_dfr(turb_files, read_parquet)%>%
  adjust_turbidity_window(
    site_oi      = "cbri",
    win_start = "2023-09-05 11:45",
    win_end   = "2023-10-10 12:45",
    dt_before = "2023-10-10 13:15",
    dt_start  = "2023-10-10 12:45"
  )%>%
  adjust_turbidity_window(
    site_oi      = "salyer",
    win_start = "2023-08-16 07:45",
    win_end   = "2023-09-19 14:15",
    dt_before = "2023-09-22 08:00",
    dt_start  = "2023-09-19 14:15"
  )%>%
  adjust_turbidity_window(
    site_oi      = "cottonwood_virridy",
    win_start = "2023-10-10 13:45",
    win_end   = "2023-12-06 11:30",
    dt_before = "2023-10-10 16:45",
    dt_start  = "2023-10-10 11:45"
  )%>%
  adjust_turbidity_window(
    site_oi      = "archery",
    win_start = "2023-04-28 15:30",
    win_end   = "2023-05-08 07:00",
    dt_before = "2023-04-28 15:15",
    dt_start  = "2023-04-28 18:00"
  )%>%
  mutate(visited = ifelse(auto_flag == "site visit window" | auto_flag == "site visit", TRUE, FALSE))

# plot_data <- turb_data%>%filter(site == "joei" & mean_pre_cal <= 100)
# 
# ggplotly(ggplot(plot_data,
#                 aes(x = DT_round, y = adj_turb))+
#            geom_point(aes(y = mean_pre_cal), color = "grey", alpha = 0.5)+
#   geom_point(aes(color = user_flag))+
#     geom_line(aes(y = mean_drift_trans))+
#     geom_vline(xintercept = plot_data$DT_round[plot_data$visited == TRUE], color = "red", linetype = "dashed")+
#   labs(title = "Turbidity Data with Adjustments",
#        x = "Date",
#        y = "Turbidity (NTU)",
#        color = "Turbidity Corrected")+
#   facet_wrap(~site, scales = "free_y"))
final_turb_data <- turb_data%>%
  select(DT_round, DT_join, site, parameter,units, 
         value_raw = mean_pre_cal, 
         value_post_cal_corr = mean, 
         value_post_ver = mean_analysis, 
         value_post_drift_corr = adj_turb,   
         auto_qc_flag = auto_flag, mal_flag, 
         user_qc_flag = user_flag, qc_status = final_status)
```

## Organizing Data for pub

We will be creating two datasets here:

- A "raw" dataset that includes all the original data with flags and all "versions" of the data:

  - no manual corrections applied, calibration correction, post verification, and post drift correction.

  - This is for posterity and to have a record of what the original data looked like before and after manual verification/corrections.

- A "cleaned" dataset that includes only the final value (post drift correction if it exists, post verification if not) and a column indicating what modifications were made to the data (manual exclusion, calibration correction, drift correction, etc).

Both datasets will be uploaded to HydroShare

```{r}
fdom_files <- list.files(path = 'data/raw/sensor/manual_data_verification/2023_cycle/post_verification/', pattern = "*.parquet", full.names = TRUE)%>%grep(pattern = "FDOM", x = ., ignore.case = TRUE,value = T)

drift_corrected_data <- map_dfr(fdom_files, read_parquet)%>%
  select(DT_round, DT_join, site, parameter,units, 
         value_raw = mean_pre_cal, 
         value_post_cal_corr = mean, 
         value_post_ver = mean_analysis, 
         value_post_drift_corr = mean_drift_trans,   
         auto_qc_flag = auto_flag, mal_flag, 
         user_qc_flag = user_flag, qc_status = final_status)%>%
  bind_rows(final_turb_data)%>%
  split(f = list(.$site, .$parameter), sep = "-")%>% 
  keep(~ nrow(.x) > 0) #remove any empty datasets

non_drift_corrected_files <- list.files(path = 'data/raw/sensor/manual_data_verification/2023_cycle/in_progress/verified_directory/', pattern = "*.parquet", full.names = TRUE)%>% grep(pattern = "Depth", x = ., ignore.case = TRUE,value = T,invert = T)

non_drift_corrected_data <- map_dfr(non_drift_corrected_files, read_parquet)%>%
  mutate( mean_analysis = ifelse(verification_status %in% c("FLAGGED", "PASS"), mean, NA))%>%
  select(DT_round, DT_join, site, parameter,units,
         value_raw = mean_pre_cal, 
         value_post_cal_corr = mean, 
         value_post_ver = mean_analysis,
         auto_qc_flag = auto_flag, 
         mal_flag, 
         user_qc_flag = user_flag,
         qc_status = final_status)%>%
  #adding in the final depth data from above
  bind_rows(final_depth_data)%>% 
  split(f = list(.$site, .$parameter), sep = "-")

#only keep non drift corrected site parameters if they are not in the drift corrected data (i.e. if they haven't been drift corrected yet)
all_datasets <- c(non_drift_corrected_data[!names(non_drift_corrected_data) %in% names(drift_corrected_data)], drift_corrected_data)%>%
  keep(~ nrow(.x) > 0) #remove any empty datasets

#units lookup table
#--------------------------#
#DOUBLE CHECK IF NEW PARAMETERS ARE ADDED OR UNITS CHANGE
#--------------------------#
units_table <- tibble(
  parameter = c("Chl-a Fluorescence", "Depth", "DO", "ORP", "Specific Conductivity", "Temperature", "FDOM Fluorescence", "Turbidity"),
  units = c("RFU", "m", "mg/L", "V", "µS/cm", "C", "RFU", "NTU")
)

raw_dataset <- all_datasets%>%
  bind_rows()%>%
  #exclude springcreek and boxcreek since data quality is a bit suspect
  filter(!(site %in% c("springcreek", "boxcreek")))%>%
  mutate(
    #keep turb and fDOM from archery_virridy site from 9/2023 and 12/23
    is_archery_window = str_detect(site, "archery_virridy") &
      between(DT_round, ymd_hms("2023-09-01 00:00:00", tz = "MST"), ymd_hms("2023-12-31 23:59:59", tz = "MST")) & 
      parameter %in% c("Turbidity", "FDOM Fluorescence"),
    #keep fDOM and Chl-a from riverbend and cottonwood virridy sites from 9/2023 and 12/23
    is_river_cotton_window = str_detect(site, "riverbend_virridy|cottonwood_virridy") &
      between(DT_round, ymd_hms("2023-09-01 00:00:00", tz = "MST"), ymd_hms("2023-05-31 23:59:59", tz = "MST")) & 
      parameter %in% c("FDOM Fluorescence", "Chl-a Fluorescence"),
    site = if_else(is_archery_window | is_river_cotton_window, str_remove(site, "_virridy"), site)
  ) %>%
  filter(
    site %nin% c("archery_virridy", "riverbend_virridy", "cottonwood_virridy")
  )%>%
  mutate(DateTime_UTC = with_tz(DT_round, tzone = "UTC"),#convert to UTC
         #Salyer had a miscalibration of the pH sensor that was not corrected in the calibration correction tool. This is a temporary fix to correct the pH values for this site and parameter.
         is_window_one = DateTime_UTC >= ymd_hm("2023-09-22 20:30", tz = "UTC") & DateTime_UTC <= ymd_hm("2023-10-12 17:30", tz = "UTC") & site == "salyer" & parameter == "pH"& !is.na(value_post_ver),
           value_post_cal_corr = case_when(
             is_window_one ~ value_post_cal_corr - 0.25,
             TRUE            ~ value_post_cal_corr
           ), 
         value_post_ver = case_when(
           is_window_one ~ value_post_ver - 0.25,
           TRUE            ~ value_post_ver
         ))%>%
  # if value_post_drift corr is NA, then fill with value_post_ver (this means that the data was verified but not drift corrected, so the post verification value is the most "final" version of the data we have for those instances) This is for parameters other than FDOM and Turbidity
  mutate(value_post_drift_corr = if_else(is.na(value_post_drift_corr), value_post_ver, value_post_drift_corr),
         #round all values to 3 decimal places for consistency with HydroShare upload
         value_raw = round(value_raw, 3),
         value_post_cal_corr = round(value_post_cal_corr, 3),
         value_post_ver = round(value_post_ver, 3),
         value_post_drift_corr = round(value_post_drift_corr, 3),
  # Descriptions to the modifications column
    modifications = case_when(
      is.na(value_post_ver) & !is.na(value_raw) ~ "manual data exclusion",
      value_raw == value_post_cal_corr & !is.na(value_post_ver) & value_post_ver == value_post_drift_corr & grepl("drift", user_qc_flag) ~ "non-resolved drift",
      value_raw == value_post_cal_corr & value_post_ver != value_post_drift_corr ~ "drift correction",
      value_raw != value_post_cal_corr & value_post_ver == value_post_drift_corr ~ "calibration correction",
      value_raw != value_post_cal_corr & value_post_ver != value_post_drift_corr ~ "calibration correction;drift correction",
      TRUE ~ NA_character_
     ), 
    value_post_drift_corr = if_else( !is.na(modifications) & modifications == "non-resolved drift" & parameter != "Chl-a Fluorescence", NA, value_post_drift_corr))%>%
  select(-units)%>%
  left_join(units_table, by = "parameter")%>%
  select(DateTime_UTC,
         site, parameter, units, 
         #Actual data
         value_raw, 
         value_post_cal_corr,
         value_post_ver,
         value_post_drift_corr, 
         #QC Info
         auto_qc_flag, mal_flag, user_qc_flag, qc_status, modifications)%>%
  #ensure turbidity is > 0 for value_post_drift_corr, value_post_ver, value_post_cal_corr, and value_raw. If not, set to 0.1
  mutate(value_post_drift_corr = if_else(parameter == "Turbidity" & value_post_drift_corr <= 0.1, 0.1, value_post_drift_corr),
         value_post_ver = if_else(parameter == "Turbidity" & value_post_ver <= 0.1, 0.1, value_post_ver),
         value_post_cal_corr = if_else(parameter == "Turbidity" & value_post_cal_corr <= 0.1, 0.1, value_post_cal_corr))
    
final_dataset <- raw_dataset %>%
  select(DateTime_UTC, site, parameter, units, value_raw, value_final = value_post_drift_corr,user_qc_flag, modifications)

```

## Plot to Double Check

```{r}
plots <- map(unique(final_dataset$site), function(site_name) {
  
  ggplot(final_dataset %>% 
           filter(site == site_name )%>%
           #remove turbidity values greater than 1000 NTU since they are likely erroneous and will skew the y-axis
           filter(!(parameter == "Turbidity" & value_final > 1000)), 
         aes(x = DateTime_UTC, y = value_final, color = modifications)) +
    geom_point(aes(y = value_raw),size = 0.5, alpha = 0.1) + # add raw data points for comparison
    geom_point() +
    facet_wrap(~parameter, scales = "free_y") +
    labs(title = site_name,
         x = "Date",
         y = "Value",
         color = "QC Status")
})%>%
  set_names(unique(final_dataset$site))
#Best to run each site individually
#walk(plots, print)
```

### One specific site-parameter

```{r}
additional_site <- final_dataset %>% filter(site == "elc" & parameter == "Turbidity")

#Check out a specific parameter for a specific site
ggplotly(ggplot(final_dataset %>% 
         filter(site == "salyer" & parameter == "Turbidity"), 
       aes(x = DateTime_UTC, y = value_final, color = modifications)) +
  #geom_point(aes(y = value_raw),size = 0.5, alpha = 0.1) + # add raw data points for comparison
  geom_point()+
    geom_line(data = additional_site, aes(x = DateTime_UTC, y = value_final), color = "black")+
  facet_wrap(~parameter, scales = "free_y"))

```

## Save Data

For each year we will save a csv and parquet version of each file

```{r}
year <- "2023"

#check if the folder exists, if not create it
if (!dir.exists(here( "data/collated/sensor/hydroshare_pub", year))) {
  dir.create(here( "data/collated/sensor/hydroshare_pub", year), recursive = TRUE)
}

cleaned_file_name <- here( "data/collated/sensor/hydroshare_pub", year, paste0("CSU_ROSS_", year,"_cleaned.csv"))

raw_file_name <- here( "data/collated/sensor/hydroshare_pub", year, paste0("CSU_ROSS_", year,"_raw.csv"))

write_csv(final_dataset, cleaned_file_name)

write_csv(raw_dataset, raw_file_name)
# Repeat for parquet files
cleaned_file_name_parquet <- here( "data/collated/sensor/hydroshare_pub", year, paste0("CSU_ROSS_", year,"_cleaned.parquet"))

raw_file_name_parquet <- here( "data/collated/sensor/hydroshare_pub", year, paste0("CSU_ROSS_", year,"_raw.parquet"))

write_parquet(final_dataset, cleaned_file_name_parquet)

write_parquet(raw_dataset, raw_file_name_parquet)
```

## Field Notes

Since we transitioned from old field notes to our mWater system, I am taking a couple extra steps to clean up/standardize the field notes for downstream use

```{r}

old_field_notes <- read_excel(here("data","raw","field_notes","sensor_field_notes.xlsx")) %>% 
  mutate(DT = (paste0(date, " ", start_time_mst))) %>%
  mutate(DT = ymd_hm(DT) + hours(7)) %>%
  arrange(DT) %>%
  mutate(DT_round = round_date(DT, "15 minutes")) %>%
  mutate(DT_round = with_tz(DT_round, tzone = "UTC"),
         last_site_visit = with_tz(DT_round, tzone = "UTC"),
         DT_join = as.character(DT_round),
         sonde_moved = NA,
         sonde_employed = case_when(is.na(sensor_deployed) & is.na(sensor_pulled) ~ NA,
                                    sensor_deployed == "x" ~ 0, 
                                    sensor_pulled == "x" ~ 1), 
         site = ifelse(site == "rist", "bellvue", site)) %>%
  ross.wq.tools::fix_site_names()%>%
  filter(!(site %in% c("springcreek", "boxcreek", "mountaincampus")))%>%
  filter(DT_round >= ymd("2022-12-31"))%>%
  rename(DateTime_UTC = DT_round) %>%
  select(-any_of(c("sensors_cleaned", "wiper_working", "photos_downloaded","DT_join", "cal_report_collected", "sensor_swapped_notes", "end_dt", "last_site_visit", "date","sensor_deployed", "sensor_pulled")))

#mini function to clean up the cals_performed column in the field notes data. This will standardize the values to a consistent format and remove any variations of "none" or NA.
clean_cals <- function(text) {
  # 1. Clean NA and "none" variations directly inside the function
  if (is.na(text) || str_to_lower(str_trim(text)) %in% c("none", "")) {
    return("None")
  }
  
  cals <- character()
  
  # 2. Detect parameters via regex keyword search
  if (str_detect(text, regex("turb", ignore_case = TRUE)))     cals <- c(cals, "Turbidity")
  if (str_detect(text, regex("r?do", ignore_case = TRUE)))      cals <- c(cals, "DO")
  if (str_detect(text, regex("cond", ignore_case = TRUE)))     cals <- c(cals, "Conductivity")
  if (str_detect(text, regex("\\bph\\b", ignore_case = TRUE))) cals <- c(cals, "pH")
  if (str_detect(text, regex("orp", ignore_case = TRUE)))      cals <- c(cals, "ORP")
  if (str_detect(text, regex("level", ignore_case = TRUE)))    cals <- c(cals, "Level")
  
  # Return original string if no known keywords match (e.g., custom notes)
  if (length(cals) == 0) return(text) 
  
  # 3. Sort alphabetically and combine into a clean, unified string
  paste(sort(unique(cals)), collapse = ", ")
}

sensor_field_notes <- grab_mWater_sensor_notes(load_mWater())%>% 
  filter(year(DT_round) == year & !(site %in% c("springcreek", "boxcreek")))%>%
  rename(DateTime_UTC = DT_round) %>%
  select(-photos_downloaded,-DT_join, -cal_report_collected, -sensor_swapped_notes, -end_dt, -last_site_visit, -date,-sensor_deployed, -sensor_pulled)%>%
  bind_rows(formatted_notes)%>%
    #Clean up crew to just initials and remove any extra text
  mutate(crew = crew %>%
    str_replace_all("Other \\(please specify\\)", "Other") %>%
    str_replace_all("([A-Z])[a-z]+", "\\1") %>%
    str_replace_all("(?<=[A-Z])\\s+(?=[A-Z])", "")%>%
    str_replace_all("sam struthers", "SS")%>%
    str_replace_all("sam", "S")%>%
    str_replace_all("and", ",")%>% 
    str_replace_all(" ", "")%>%
    str_replace_all("_", "")
  )%>%
  #replace any character columns that are "NA" or "na" with NA
  mutate(across(where(is.character), ~ na_if(., "NA")))%>%
  mutate(across(where(is.character), ~ na_if(., "na")))%>%
  mutate(cals_performed = map_chr(cals_performed, clean_cals))%>%
  mutate(ph_junction_replaced = ifelse(str_to_title(ph_junction_replaced) == "No"| is.na(ph_junction_replaced), "No", ph_junction_replaced),
         rdo_cap_replaced = ifelse(str_to_title(rdo_cap_replaced) == "No" | is.na(rdo_cap_replaced), "No", rdo_cap_replaced))%>%
   # across(c(ph_junction_replaced,rdo_cap_replaced )), ~ ifelse(str_to_title(.,) == "No", "No", .x))%>%
    select(any_of(c("site",  "crew","DateTime_UTC", "sonde_employed", "sensors_cleaned", "wiper_working",
                  "rdo_cap_condition", "rdo_cap_replaced", "ph_junction_replaced", "cals_performed",
                  "sonde_moved", "sensor malfunction", "visit_type", "start_DT", "visit_comments", 
                  "rdo_pre_clean",   "rdo_post_clean", "rdo_post_cal",
                  "cond_pre_clean", "cond_post_clean", "cond_post_cal",
                  "ph_pre_clean", "ph_post_clean", "ph_post_cal",
                  "orp_pre_clean", "orp_post_clean", "orp_post_cal",
                  "turb_pre_clean", "turb_post_clean", "turb_post_cal",
                  "fdom_pre_clean", "fdom_post_clean", "fdom_post_cal",
                  "chla_pre_clean", "chla_post_clean", "chla_post_cal")))%>%
  fix_site_names()

field_notes_file <- here("data/collated/sensor/hydroshare_pub", year, paste0("CSU_ROSS_sensor_field_notes_", year, ".csv"))

write_csv(sensor_field_notes, field_notes_file)

```

## Location Metadata

Here we will load and double check the location metadata file created by Sam Struthers and verified by Katie Willi

```{r}
 location_meta <- read_csv("data/collated/sensor/hydroshare_pub/CSU_ROSS_sonde_location_metadata.csv",
                           show_col_types = FALSE)%>%
  mutate(data_start_date = as.POSIXct(data_start_date, tz = "UTC", format = "%m/%d/%y"),
         data_end_date =  if_else( is.na(data_end_date), 
                                   ymd("2025-11-17"),
                                   as.POSIXct(data_end_date, tz = "UTC", format = "%m/%d/%y")))%>%

  mutate(min_year = year(data_start_date),
         max_year = year(data_end_date), 
          year_range = paste0(min_year, "-", max_year)) 

#update as needed
#write_csv(location_meta, here("data/collated/sensor/hydroshare_pub/CSU_ROSS_sonde_location_metadata.csv"))
# 
# tmap_mode("view")
# tm_basemap("OpenStreetMap") +
#   tm_shape(location_meta %>% st_as_sf(coords = c("longitude", "latitude"), crs = 4326)) +
#   #color is SiteID and shape is year_range
#     tm_symbols(col = "SiteID", shape = "year_range", size = 0.5, border.col = "black")
# 


```







