---
title: "Process_2019_2021"
author: "Katie Willi"
date: "2025-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the workflow for organizing old (2019-2021) Poudre Sonde Network data. This data was collected before our current protocol was produced, and therefore requires a slightly different workflow for performing our auto-QAQC pipeline. 

## 1) Packages & parallel configuration

```{r}
# load required packages
package_loader <- function(x) {
  if (x %in% installed.packages()) {
    suppressMessages({
      library(x, character.only = TRUE)
    })
  } else {
    suppressMessages({
      install.packages(x)
      library(x, character.only = TRUE)
    })
  }
}
non_cran_packages <- c("rossyndicate/ross.wq.tools")
invisible(
  lapply(non_cran_packages, function(x) {
    pack_name <- unlist(strsplit(x, "/"))[2] #get package name (no GH username)

    if (pack_name %in% installed.packages()) {
      suppressMessages({
        library(pack_name, character.only = TRUE)
      })
    } else {
      suppressMessages({
        devtools::install_github(x)
        library(pack_name, character.only = TRUE)
      })
    }
  })
)
# load all required packages
invisible(
  lapply(c("arrow", "data.table", "httr2", "tidyverse", "lubridate", "zoo", "ggpubr",
           "padr", "stats", "RcppRoll", "yaml", "here", "furrr", "ross.wq.tools"),
         package_loader)
)
```

## 2) Thresholds, credentials, and field notes data

```{r}
# Read in the threshold data first
sensor_thresholds <- read_yaml(here("data","derived","auto_qaqc_files", "thresholds", "sensor_spec_thresholds.yml"))
season_thresholds <- read_csv(here("data","derived","auto_qaqc_files", "thresholds","seasonal_thresholds20260304-T175010Z.csv"), show_col_types = FALSE) %>%
  fix_site_names()

# Configure your credentials files
mwater_creds_file <- here("creds","mWaterCreds.yml")
mWater_creds <- read_yaml(mwater_creds_file)

# pull field data from mWater API
mWater_data <- load_mWater(creds = mWater_creds)

# grab field notes with proper timezone handling
new_field_notes <- grab_mWater_sensor_notes(mWater_api_data = mWater_data) %>%
  mutate(DT_round = with_tz(DT_round, tzone = "UTC"),
         last_site_visit = with_tz(last_site_visit, tzone = "UTC"),
         DT_join = as.character(DT_round))


old_field_notes <- readxl::read_excel(here("data","raw","field_notes","sensor_field_notes.xlsx")) %>%
  mutate(DT = (paste0(date, " ", start_time_mst))) %>%
  mutate(DT = ymd_hm(DT) + hours(7)) %>%
  arrange(DT) %>%
  mutate(DT_round = round_date(DT, "15 minutes")) %>%
  mutate(DT_round = with_tz(DT_round, tzone = "UTC"),
         last_site_visit = with_tz(DT_round, tzone = "UTC"),
         DT_join = as.character(DT_round),
         sonde_moved = NA,
         sonde_employed = case_when(is.na(sensor_deployed) & is.na(sensor_pulled) ~ NA,
                                    sensor_deployed == "x" ~ 0, 
                                    sensor_pulled == "x" ~ 1), 
         site = ifelse(site == "rist", "bellvue", site)) %>%
  ross.wq.tools::fix_site_names()%>%
  #add extra cols
  mutate(rdo_pre_clean = NA, 
         rdo_post_clean = NA,
         rdo_post_cal = NA,
         cond_pre_clean = NA,
         cond_post_clean = NA,
         cond_post_cal = NA,
         ph_pre_clean = NA,
         ph_post_clean = NA,
         ph_post_cal = NA,
         orp_pre_clean = NA,
         orp_post_clean = NA,
         orp_post_cal = NA,
         turb_pre_clean = NA,
         turb_post_clean = NA,
         turb_post_cal = NA,
         fdom_pre_clean = NA,
         fdom_post_clean = NA,
         fdom_post_cal = NA,
         chla_pre_clean = NA,
         chla_post_clean = NA,
         chla_post_cal = NA)%>%
  select(-sensors_cleaned, -wiper_working, -cal_report_collected)

all_field_notes <- bind_rows(new_field_notes, old_field_notes) 

# grab sensor malfunction records
sensor_malfunction_notes <- grab_mWater_malfunction_notes(mWater_api_data = mWater_data) %>%
  mutate(start_DT = with_tz(start_DT, tzone = "UTC"),
         end_DT = with_tz(end_DT, tzone = "UTC"))
```

## 3) Ingest & standardize raw data

```{r}
# HTML readers, one for HydroVu htmls, one for direct downloads from the AquaTROLL 600, and one for downloading from the Vulink in the field:

# from HydroVu:
hydrovu_reader <- function(file) {
  
  raw_data <- rvest::read_html(file) %>%
    rvest::html_node('table') %>%
    rvest::html_table() %>%
    slice(-1:-8) %>%
    janitor::row_to_names(row_number = 1) 
}

# from AquaTROLL 600:
troll_reader <- function(file) {
  
  raw_data <- rvest::read_html(file) %>%
    rvest::html_node('table') %>%
    rvest::html_table() %>%
    slice(-1:-25) %>%
    janitor::row_to_names(row_number = 1) 
}

# from VuLink in the field:
vulink_reader <- function(file) {
  
  raw_data <- rvest::read_html(file) %>%
    rvest::html_node('table') %>%
    rvest::html_table() %>%
    slice(-1:-31) %>%
    janitor::row_to_names(row_number = 1) 
}
```

**Ingest data from each site:**

```{r}
base_path <- here("data", "raw", "sensor", "log_download", "2018_2021", "site_organized_hydrovu")
## Rist

hydrovu_utc <- map_dfr(list.files(here(base_path, "rist"), full.names = T), hydrovu_reader) %>%
  select(DT = `Date Time`,
         Water_Temp_C = `Temperature (C)`,
         pH = `pH (pH)`,
         ORP_mV = `ORP (mV)`,
         Actual_Conductivity_µS_cm = `Actual Conductivity (µS/cm)`,
         Specific_Conductivity_µS_cm = `Specific Conductivity (µS/cm)`,
         Salinity_PSU = `Salinity (psu)`,
         TDS_ppt = `Total Dissolved Solids (ppt)`,
         DO_ppm = `DO (ppm)`,
         Perc_O2_Sat = `% Saturation O₂ (% sat)`,
         Turbidity_NTU = `Turbidity (FNU)`,
         TSS_mg_L = `Total Suspended Solids (mg/L)`,
         #Chla_RFU = Chl.a.Fluorescence..RFU.,
         #Chla_µg_L = Chl.a.Concentration..µg.L.,
         Pressure_PSI = `Pressure (psi)`,
         Depth_ft = `Depth (ft)`,
         # Elevation_m = `Level: Elevation (m)`,
         Depth_to_Water_ft = `Level: Depth to Water (ft)`) %>%
  mutate(DT = as_datetime(DT, tz = "UTC")) 

rist <- hydrovu_utc %>%
  mutate(year = lubridate::year(DT)) %>%
  mutate(date = as_date(substr(DT, 1, 10))) %>%
  distinct(.keep_all = TRUE) %>%
  # Force time zones appropriately. MST to UTC:
  mutate(DT = case_when(year %in% c(2019, 2020) ~ DT + hours(7),
                        TRUE ~ DT)) %>% 
  mutate(sensor = "rist",
         comments = NA)

rm(hydrovu_utc)

## Legacy

hydrovu_utc <- map_dfr(list.files(here(base_path,"legacy"), pattern = "hydrovu", full.names = T), hydrovu_reader) %>%
  select(DT = `Date Time`,
         Water_Temp_C = `Temperature (C)`,
         pH = `pH (pH)`,
         ORP_mV = `ORP (mV)`,
         Actual_Conductivity_µS_cm = `Actual Conductivity (µS/cm)`,
         Specific_Conductivity_µS_cm = `Specific Conductivity (µS/cm)`,
         Salinity_PSU = `Salinity (psu)`,
         TDS_ppt = `Total Dissolved Solids (ppt)`,
         DO_ppm = `DO (ppm)`,
         Perc_O2_Sat = `% Saturation O₂ (% sat)`,
         Turbidity_NTU = `Turbidity (FNU)`,
         TSS_mg_L = `Total Suspended Solids (mg/L)`,
         #Chla_RFU = Chl.a.Fluorescence..RFU.,
         #Chla_µg_L = Chl.a.Concentration..µg.L.,
         Pressure_PSI = `Pressure (psi)`,
         Depth_ft = `Depth (ft)`,
         # Elevation_m = `Level: Elevation (m)`,
         Depth_to_Water_ft = `Level: Depth to Water (ft)`) %>%
  mutate(DT = as_datetime(DT, tz = "UTC"))

troll_mdt <- map_dfr(list.files(here(base_path, "legacy"), full.names = T, pattern = "troll_mdt"), troll_reader)
names(troll_mdt) <- make.names(gsub("[[:digit:]]", "", names(troll_mdt)), unique = T)
troll_mdt <- troll_mdt %>%
  select(DT = Date.Time,
         Water_Temp_C = Temperature...C....,
         pH = pH..pH....,
         ORP_mV = ORP..mV....,
         Actual_Conductivity_µS_cm = Actual.Conductivity..µS.cm....,
         Specific_Conductivity_µS_cm = Specific.Conductivity..µS.cm....,
         Salinity_PSU = Salinity..PSU....,
         TDS_ppt = Total.Dissolved.Solids..ppt....,
         DO_ppm = RDO.Concentration..mg.L....,
         Perc_O2_Sat = RDO.Saturation...Sat....,
         Turbidity_NTU = Turbidity..NTU....,
         TSS_mg_L = Total.Suspended.Solids..mg.L....,
         Pressure_PSI = Pressure..psi....,
         Depth_ft = Depth..ft....) %>%
  mutate(DT_local = force_tz(ymd_hms(DT), tzone = "America/Denver"),
         DT = with_tz(DT_local, tzone = "UTC")) %>%
  filter(DT_local >= "2021-07-25 00:00:00" & DT_local <="2021-08-25 23:00:00") 

vulink_mdt <-  map_dfr(list.files(here(base_path, "legacy"), full.names = T, pattern = "vulink_mdt"), vulink_reader) %>% select(-2:-4)
names(vulink_mdt) <- make.names(gsub("[[:digit:]]", "", names(vulink_mdt)), unique = T)
vulink_mdt <- vulink_mdt %>%
  select(DT = Date.Time,
         Water_Temp_C = Temperature...C....,
         pH = pH..pH....,
         ORP_mV = ORP..mV....,
         Actual_Conductivity_µS_cm = Actual.Conductivity..µS.cm....,
         Specific_Conductivity_µS_cm = Specific.Conductivity..µS.cm....,
         Salinity_PSU = Salinity..PSU....,
         TDS_ppt = Total.Dissolved.Solids..ppt....,
         DO_ppm = RDO.Concentration..mg.L....,
         Perc_O2_Sat = RDO.Saturation...Sat....,
         Turbidity_NTU = Turbidity..NTU....,
         TSS_mg_L = Total.Suspended.Solids..mg.L....,
         Pressure_PSI = Pressure..psi....,
         Depth_ft = Depth..ft....) %>%
  mutate(DT_local = force_tz(ymd_hms(DT), tzone = "America/Denver"),
         DT = with_tz(DT_local, tzone = "UTC")) %>%
  filter(DT_local < "2021-07-25 00:00:00" | DT_local >"2021-08-25 23:00:00")

troll_mst <- map_dfr(list.files(here(base_path, "legacy"), full.names = T, pattern = "troll_mst"), troll_reader)
names(troll_mst) <- make.names(gsub("[[:digit:]]", "", names(troll_mst)), unique = T)
troll_mst <- troll_mst %>% 
  select(DT = Date.Time,
         Water_Temp_C = Temperature...C....,
         pH = pH..pH....,
         ORP_mV = ORP..mV....,
         Actual_Conductivity_µS_cm = Actual.Conductivity..µS.cm....,
         Specific_Conductivity_µS_cm = Specific.Conductivity..µS.cm....,
         Salinity_PSU = Salinity..PSU....,
         TDS_ppt = Total.Dissolved.Solids..ppt....,
         DO_ppm = RDO.Concentration..mg.L....,
         Perc_O2_Sat = RDO.Saturation...Sat....,
         Turbidity_NTU = Turbidity..NTU....,
         #TSS_mg_L = Total.Suspended.Solids..mg.L....,
         Pressure_PSI = Pressure..psi....,
         Depth_ft = Depth..ft....) %>%
  mutate(DT_local = force_tz(parse_date_time(DT, orders = c("ymd HMS", "ymd HM", "mdy HMS", "mdy HM")), tzone = "MST7MDT"),
         DT = with_tz(DT_local + hours(1), "UTC"))

legacy <- bind_rows(hydrovu_utc, troll_mdt, vulink_mdt, troll_mst) %>%
  # mutate(DT = round_date(DT, "15 minutes")) %>%
  mutate(year = lubridate::year(DT)) %>%
  mutate(date = as_date(substr(DT, 1, 10))) %>%
  distinct(.keep_all=TRUE) %>%
  # Force time zones appropriately. MDT to UTC:
  mutate(DT = case_when(year %in% c(2019, 2020) ~ DT + hours(6),
                        TRUE ~ DT)) %>% 
  mutate(sensor = "legacy", comments = NA) %>%
  select(-DT_local)

rm(hydrovu_utc, troll_mdt, troll_mst, vulink_mdt)

## Timberline

hydrovu_utc <- map_dfr(list.files(here(base_path, "timberline"), full.names = T), hydrovu_reader) %>%
  select(DT = `Date Time`,
         Water_Temp_C = `Temperature (C)`,
         pH = `pH (pH)`,
         ORP_mV = `ORP (mV)`,
         Actual_Conductivity_µS_cm = `Actual Conductivity (µS/cm)`,
         Specific_Conductivity_µS_cm = `Specific Conductivity (µS/cm)`,
         Salinity_PSU = `Salinity (psu)`,
         TDS_ppt = `Total Dissolved Solids (ppt)`,
         DO_ppm = `DO (ppm)`,
         Perc_O2_Sat = `% Saturation O₂ (% sat)`,
         Turbidity_NTU = `Turbidity (FNU)`,
         TSS_mg_L = `Total Suspended Solids (mg/L)`,
         #Chla_RFU = Chl.a.Fluorescence..RFU.,
         #Chla_µg_L = Chl.a.Concentration..µg.L.,
         Pressure_PSI = `Pressure (psi)`,
         Depth_ft = `Depth (ft)`,
         # Elevation_m = `Level: Elevation (m)`,
         Depth_to_Water_ft = `Level: Depth to Water (ft)`) %>%
  mutate(DT = as_datetime(DT, tz = "UTC")) 

timberline <- hydrovu_utc %>% 
  # mutate(DT = round_date(DT, "15 minutes")) %>%
  mutate(year = lubridate::year(DT)) %>%
  mutate(date = as_date(substr(DT, 1, 10))) %>%
  distinct(.keep_all = TRUE) %>%
  # Force time zones appropriately. MST to UTC:
  mutate(DT = if_else(year == 2020, DT + hours(7), DT)) %>% 
  mutate(sensor = "timberline",
         comments = NA)

rm(hydrovu_utc)

## Prospect
# Was not deployed in 2021 due to damage to the PVC.
hydrovu_utc <- map_dfr(list.files(here(base_path, "prospect"), full.names = T), hydrovu_reader) %>%
  select(DT = `Date Time`,
         Water_Temp_C = `Temperature (C)`,
         pH = `pH (pH)`,
         ORP_mV = `ORP (mV)`,
         Actual_Conductivity_µS_cm = `Actual Conductivity (µS/cm)`,
         Specific_Conductivity_µS_cm = `Specific Conductivity (µS/cm)`,
         Salinity_PSU = `Salinity (psu)`,
         TDS_ppt = `Total Dissolved Solids (ppt)`,
         DO_ppm = `DO (ppm)`,
         Perc_O2_Sat = `% Saturation O₂ (% sat)`,
         Turbidity_NTU = `Turbidity (FNU)`,
         TSS_mg_L = `Total Suspended Solids (mg/L)`,
         #Chla_RFU = Chl.a.Fluorescence..RFU.,
         #Chla_µg_L = Chl.a.Concentration..µg.L.,
         Pressure_PSI = `Pressure (psi)`,
         Depth_ft = `Depth (ft)`,
         # Elevation_m = `Level: Elevation (m)`,
         Depth_to_Water_ft = `Level: Depth to Water (ft)`) %>%
  mutate(DT = as_datetime(DT, tz = "UTC")) 

prospect <- hydrovu_utc %>%
  # mutate(DT = round_date(DT, "15 minutes")) %>%
  mutate(year = lubridate::year(DT)) %>%
  mutate(date = as_date(substr(DT, 1, 10))) %>%
  distinct(.keep_all=TRUE) %>%
  # Force time zones appropriately. MDT to UTC:
  mutate(DT = case_when(date < "2020-09-01" ~ DT + hours(6),
                        # MST to UTC:
                        date >= "2020-09-01" ~ DT + hours(7),
                        TRUE ~ DT)) %>% 
  mutate(sensor = "prospect",
         comments = NA)

rm(hydrovu_utc)

## ELC

hydrovu_utc <- map_dfr(list.files(here(base_path, "elc"), full.names = T, pattern = "hydrovu_utc"), hydrovu_reader) %>%
  select(DT = `Date Time`,
         Water_Temp_C = `Temperature (C)`,
         pH = `pH (pH)`,
         ORP_mV = `ORP (mV)`,
         Actual_Conductivity_µS_cm = `Actual Conductivity (µS/cm)`,
         Specific_Conductivity_µS_cm = `Specific Conductivity (µS/cm)`,
         Salinity_PSU = `Salinity (psu)`,
         TDS_ppt = `Total Dissolved Solids (ppt)`,
         DO_ppm = `DO (ppm)`,
         Perc_O2_Sat = `% Saturation O₂ (% sat)`,
         Turbidity_NTU = `Turbidity (FNU)`,
         TSS_mg_L = `Total Suspended Solids (mg/L)`,
         #Chla_RFU = Chl.a.Fluorescence..RFU.,
         #Chla_µg_L = Chl.a.Concentration..µg.L.,
         Pressure_PSI = `Pressure (psi)`,
         Depth_ft = `Depth (ft)`,
         # Elevation_m = `Level: Elevation (m)`,
         Depth_to_Water_ft = `Level: Depth to Water (ft)`) %>%
  mutate(DT = as_datetime(DT, tz = "UTC")) 

hydrovu_mst <- map_dfr(list.files(here(base_path, "elc"), full.names = T, pattern = "hydrovu_mst"), hydrovu_reader) %>%
  select(DT = `Date Time`,
         Water_Temp_C = `Temperature (C)`,
         pH = `pH (pH)`,
         ORP_mV = `ORP (mV)`,
         Actual_Conductivity_µS_cm = `Actual Conductivity (µS/cm)`,
         Specific_Conductivity_µS_cm = `Specific Conductivity (µS/cm)`,
         Salinity_PSU = `Salinity (psu)`,
         TDS_ppt = `Total Dissolved Solids (ppt)`,
         DO_ppm = `DO (ppm)`,
         Perc_O2_Sat = `% Saturation O₂ (% sat)`,
         Turbidity_NTU = `Turbidity (FNU)`,
         TSS_mg_L = `Total Suspended Solids (mg/L)`,
         #Chla_RFU = Chl.a.Fluorescence..RFU.,
         #Chla_µg_L = Chl.a.Concentration..µg.L.,
         Pressure_PSI = `Pressure (psi)`,
         Depth_ft = `Depth (ft)`) %>%
  mutate(DT_local = force_tz(parse_date_time(DT, orders = c("ymd HMS", "ymd HM", "mdy HMS", "mdy HM")), tzone = "MST7MDT"),
         DT = with_tz(DT_local + hours(1), "UTC"))

troll_mdt <- map_dfr(list.files(here(base_path, "elc"), full.names = T, pattern = "troll_mdt"), troll_reader)
names(troll_mdt) <- make.names(gsub("[[:digit:]]", "", names(troll_mdt)), unique = T)
troll_mdt <- troll_mdt %>%
  select(DT = Date.Time,
         Water_Temp_C = Temperature...C....,
         pH = pH..pH....,
         ORP_mV = ORP..mV....,
         Actual_Conductivity_µS_cm = Actual.Conductivity..µS.cm....,
         Specific_Conductivity_µS_cm = Specific.Conductivity..µS.cm....,
         Salinity_PSU = Salinity..PSU....,
         TDS_ppt = Total.Dissolved.Solids..ppt....,
         DO_ppm = RDO.Concentration..mg.L....,
         Perc_O2_Sat = RDO.Saturation...Sat....,
         Turbidity_NTU = Turbidity..NTU....,
         TSS_mg_L = Total.Suspended.Solids..mg.L....,
         Pressure_PSI = Pressure..psi....,
         Depth_ft = Depth..ft....) %>%
  mutate(DT_local = force_tz(ymd_hms(DT), tzone = "America/Denver"),
         DT = with_tz(DT_local, tzone = "UTC"))

elc <- bind_rows(hydrovu_utc, hydrovu_mst, troll_mdt) %>%
  # mutate(DT = round_date(DT, "15 minutes")) %>%
  mutate(year = lubridate::year(DT)) %>%
  mutate(date = as_date(substr(DT, 1, 10))) %>%
  distinct(.keep_all=TRUE) %>%
  mutate(DT = if_else(year == 2019, DT + hours(6), DT)) %>%
  mutate(sensor = "elc",
         comments = NA) %>%
  select(-DT_local)

rm(hydrovu_mst, hydrovu_utc, troll_mdt)

## Archery

hydrovu_utc <- map_dfr(list.files(here(base_path, "archery"), full.names = T), hydrovu_reader) %>%
  select(DT = `Date Time`,
         Water_Temp_C = `Temperature (C)`,
         pH = `pH (pH)`,
         ORP_mV = `ORP (mV)`,
         Actual_Conductivity_µS_cm = `Actual Conductivity (µS/cm)`,
         Specific_Conductivity_µS_cm = `Specific Conductivity (µS/cm)`,
         Salinity_PSU = `Salinity (psu)`,
         TDS_ppt = `Total Dissolved Solids (ppt)`,
         DO_ppm = `DO (ppm)`,
         Perc_O2_Sat = `% Saturation O₂ (% sat)`,
         Turbidity_NTU = `Turbidity (FNU)`,
         TSS_mg_L = `Total Suspended Solids (mg/L)`,
         Chla_RFU = `Chl-a Fluorescence (RFU)`,
         Chla_µg_L = `Chl-a Concentration (µg/L)`,
         Pressure_PSI = `Pressure (psi)`,
         Depth_ft = `Depth (ft)`,
         # Elevation_m = `Level: Elevation (m)`,
         Depth_to_Water_ft = `Level: Depth to Water (ft)`) %>%
  mutate(DT = as_datetime(DT, tz = "UTC")) 

archery <- hydrovu_utc %>%
  # mutate(DT = round_date(DT, "15 minutes")) %>%
  mutate(year = lubridate::year(DT)) %>%
  mutate(date = as_date(substr(DT, 1, 10))) %>%
  distinct(.keep_all = TRUE) %>%
  # Force time zones appropriately. MDT to UTC:
  mutate(DT = case_when(year == 2020 ~ DT + hours(6),
                        # MST to UTC:
                        year == 2019 ~ DT + hours(7),
                        TRUE ~ DT)) %>%
  mutate(sensor = "archery",
         comments = case_when(DT == "2019-09-30 03:45:00" ~ "sensor deployed",
                              DT == "2020-01-06 21:45:00" ~ "rental sensor battery dead",
                              DT == "2020-01-10 08:45:00" ~ "rental sensor battery replaced",
                              DT == "2020-04-04 18:00:00" ~ "rental sensor battery dead",
                              DT == "2020-04-17 11:15:00" ~ "rental sensor battery replaced",
                              DT == "2020-04-24 23:00:00" ~ "rental sensor stopped working?",
                              DT == "2020-04-25 15:45:00" ~ "rental sensor swapped with new sensor",
                              DT == "2021-08-17 09:00:00" ~ "sensor cleaned",
                              DT == "2021-08-17 09:15:00" ~ "sensor re-deployed",
                              DT == "2021-09-19 05:15:00" ~ "sensor stopped working - overheating?",
                              DT == "2021-09-25 09:30:00" ~ "sensor removed from field",
                              DT == "2021-10-04 09:00:00" ~ "sensor re-deployed",
                              DT == "2021-10-05 07:45:00" ~ "sensor checked, then re-deployed",
                              DT == "2021-12-04 09:00:00" ~ "sensor pulled for season"))

rm(hydrovu_utc)

old_raw <- bind_rows(rist, legacy, timberline, prospect, elc, archery) %>%
  mutate(sensor = ifelse(sensor == "rist", "tamasag", sensor))

# rm(archery, elc, legacy, prospect, rist, timberline)

# what we want to try to "recreate" as far as formatting:
# raw_example <- read_parquet("data/raw/sensor/manual_data_verification/2024_cycle/hydro_vu_pull/raw/archery_virridy2025-01-01_0659.parquet", as_data_frame = TRUE)
```

**Standardization rules**

-   Long format by variable

-   Map old variable names → canonical `parameter` and `units`

-   Convert units (ft→m, mV→V, DO ppm→mg/L)

-   Common columns: `site`, `id`, `name`, `timestamp`, `parameter`, `value`, `units`

```{r}
raw_data_standardized <- old_raw %>%
  pivot_longer(-c(DT, year, date, sensor, comments), values_to = "value", names_to = "var") %>%
  filter(!var %in% c("Actual_Conductivity_µS_cm", "Salinity_PSU", "TDS_ppt", "Perc_O2_Sat", "TSS_mg_L",
                     "Pressure_PSI", "Elevation_m", "Depth_to_Water_ft", "Chla_µg_L")) %>%
  rename(site = sensor,
         timestamp = DT) %>%
  mutate(value = as.numeric(value),
         parameter = case_when(var == "Water_Temp_C" ~ "Temperature",
                               var == "pH" ~ "pH",
                               var == "ORP_mV" ~ "ORP",
                               var == "Specific_Conductivity_µS_cm" ~ "Specific Conductivity",
                               var == "DO_ppm" ~ "DO",
                               var ==  "Turbidity_NTU" ~ "Turbidity",
                               var == "Depth_ft" ~ "Depth",
                               var == "Chla_RFU" ~ "Chl-a Fluorescence"),
         # Original units
         original_unit = case_when(var == "Water_Temp_C" ~ "C",
                                   var == "pH" ~ "pH",
                                   var == "ORP_mV" ~ "mV",
                                   var == "Specific_Conductivity_µS_cm" ~ "µS/cm",
                                   var == "DO_ppm" ~ "ppm",
                                   var == "Turbidity_NTU" ~ "NTU",
                                   var == "Depth_ft" ~ "ft",
                                   var == "Chla_RFU" ~ "RFU"),
         # Target units
         units = case_when(var == "Water_Temp_C" ~ "C",
                           var == "pH" ~ "pH",
                           var == "ORP_mV" ~ "V",
                           var == "Specific_Conductivity_µS_cm" ~ "µS/cm",
                           var == "DO_ppm" ~ "mg/L",
                           var == "Turbidity_NTU" ~ "NTU",
                           var == "Depth_ft" ~ "m",
                           var == "Chla_RFU" ~ "RFU"),
         # Convert values to new units
         value = case_when(var == "DO_ppm" ~ value,  # ppm is equivalent to mg/L for dissolved oxygen
                           var == "Depth_ft" ~ value * 0.3048,  # feet to meters
                           var == "ORP_mV" ~ value / 1000,  # millivolts to volts
                           TRUE ~ value),
         id = site,
         name = site) %>%
  select(site, id, name, timestamp, parameter, value, units) %>%
  filter(!is.na(value)) %>%
  split(f = list(.$site), sep = "-")
```


## 4) Save site-year raw data (Parquet)

```{r}
# # Split by site and year
raw_data_site_year <- bind_rows(raw_data_standardized) %>%
  distinct(.keep_all = TRUE) %>%
  mutate(year = year(timestamp)) %>%
  split(f = list(.$site, .$year), sep = "_", drop = TRUE)



# Save files
iwalk(raw_data_site_year, ~ {
  latest_ts <- format(max(.x$timestamp, na.rm = TRUE), "%Y-%m-%d_%H%M")
  site_name <- str_split(.y, "_")[[1]][1]  # Extract just the site name (first part)
  dir_path <- here("data", "raw", "sensor", "manual_data_verification", paste0(year(latest_ts), "_cycle"), "hydro_vu_pull", "raw")
  if(!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  filename <- here(dir_path , paste0(tolower(site_name), "_", latest_ts, ".parquet"))
  .x %>% select(-year) %>% write_parquet(filename)
  cat("Saved:", filename, "\n")
})
```

## 5) Tidy, join field notes, summarize

There are, in fact, no field notes during this time frame. But we are running the data through the add_field_notes function to ensure all columns etc. are preserved for future QAQC steps.

```{r}
# tidy raw data (default 15-minute intervals)
tidy_data <-  raw_data_standardized %>%
  bind_rows() %>%
  mutate(DT_round = round_date(timestamp, "15 minutes")) %>%
  mutate(DT_join = as.character(paste(DT_round))) %>%
  ross.wq.tools::fix_site_names(., site_col = "site") %>%
  select(DT_round, DT_join, site, parameter, value, units) %>%
  distinct(.keep_all = TRUE) %>%
  modify_if(., is.numeric, ~ ifelse(is.nan(.x) | is.infinite(.x), NA, .x)) %>%
  split(f = list(.$site, .$parameter), sep = "-") %>%
  keep(~nrow(.) > 0) %>%
  future_map(~tidy_api_data(api_data = .), .progress = TRUE) %>%
  keep(~!is.null(.))

# add field notes to tidied data (there aren't any for these years)
combined_data <- tidy_data %>%
  future_map(~add_field_notes(df = ., notes = all_field_notes), .progress = TRUE)

# generate summary statistics
summarized_data <- combined_data %>%
  map(~generate_summary_statistics(.)) %>%
  # extra step to get rid of NaN and Inf values
  map(~ {modify_if(.x, is.numeric, ~ ifelse(is.nan(.x) | is.infinite(.x), NA, .x))})

```

Visualize all the data:
```{r}
plotz <- map(names(summarized_data), ~ {
  ggplot(data = summarized_data[[.x]]) +
    geom_line(aes(x = DT_round, y = mean), na.rm = TRUE) +
    labs(
      title = .x,
      x = "Date/Time",
      y = "Mean Value"
    ) +
    theme_minimal()
})

walk(plotz, print)
```


## 6) Single-sensor flags

```{r}
# Chunk the data for furrr
summarized_data_chunks <- split(1:length(summarized_data),
                                ceiling(seq_along(1:length(summarized_data))/10))
# Flag data...
# Single parameter flags

# Process the chunk in parallel
single_sensor_flags <- summarized_data %>%
  map(
    function(data) {
      flagged_data <- data %>%
        data.table(.) %>%
        # flag field visits
        # add_field_flag(df = .) %>%
        # flag missing data
        add_na_flag(df = .) %>%
        # flag DO noise
        find_do_noise(df = .) %>%
        # flag repeating values
        add_repeat_flag(df = .) %>%
        # find times when sonde was moved up/down in housing
        # add_depth_shift_flag(df = ., level_shift_table =  all_field_notes, post2024 = FALSE) %>%
        # find instances of sensor drift (FDOM, Chl-a, Turbidity only)
        add_drift_flag(df = .)
      
      if (unique(data$parameter) %in% names(sensor_thresholds)) {
        # flag instances outside the spec range
        flagged_data <- flagged_data %>%
          data.table(.) %>%
          add_spec_flag(df = ., spec_table = sensor_thresholds)
      }
      
      if (unique(data$parameter) %in% unique(season_thresholds$parameter)) {
        # flag instances outside the spec range
        flagged_data <- flagged_data %>%
          data.table(.) %>%
          add_seasonal_flag(df = ., threshold_table = season_thresholds)
      }
      
      flagged_data <- flagged_data %>%
        data.table(.)
      
      return(flagged_data)
    }
  )
```

## 7) Intrasensor flags (by site)

```{r}
# Intrasensor flags
intrasensor_flags <- single_sensor_flags %>%
  rbindlist(fill = TRUE) %>%
  split(by = "site")

# Chunk the data for furrr
intrasensor_data_chunks <- split(1:length(intrasensor_flags),
                                 ceiling(seq_along(1:length(intrasensor_flags))/3))

intrasensor_flags_list <- list()
for (chunk_idx in seq_along(intrasensor_data_chunks)) {
  message("\n=== Processing chunk ", chunk_idx, " of ", length(intrasensor_data_chunks), " ===")
  
  # Get the indices for this chunk
  indices <- intrasensor_data_chunks[[chunk_idx]]
  chunk_data <- intrasensor_flags[indices]
  # Process the chunk in parallel
  chunk_results <- chunk_data %>%
    map(
      function(data) {
        # A chunk is a site df
        flagged_data <- data %>%
          data.table() %>%
          # flag times when water was below freezing
          add_frozen_flag(.) %>%
          # overflagging correction. remove slope violation flag if it occurs concurrently
          # with temp or depth
          intersensor_check(.) %>%
          # add sonde burial. If DO is noise is long-term, likely burial:
          add_burial_flag(.) %>%
          # flag times when sonde was unsubmerged
          add_unsubmerged_flag(.)
        
        return(flagged_data)
      }, .progress = TRUE
    ) %>%
    rbindlist(fill = TRUE) %>%
    # lil' cleanup of flag column contents
    dplyr::mutate(flag = ifelse(flag == "", NA, flag)) %>%
    # transform back to site-parameter dfs
    split(f = list(.$site, .$parameter), sep = "-") %>%
    purrr::discard(~ nrow(.) == 0) %>%
    # Add in KNOWN instances of sensor malfunction
    map(~add_malfunction_flag(df = ., malfunction_records = sensor_malfunction_notes))
  
  # Add chunk to list
  intrasensor_flags_list <- c(intrasensor_flags_list, chunk_results)
  
  if (chunk_idx < length(intrasensor_data_chunks)) {
    message("Taking a short break before next chunk...")
    gc()
    Sys.sleep(0.1)
  }
}
```

## 8) Simple network consistency check

We apply a lightweight, order-aware comparison (upstream/downstream) to remove flags that are likely true environmental events observed across neighboring sites in a ±2-hour window. The site order defaults to CSU/FCW segments, with overrides for alternative networks.

```{r}
#get site order for original PWQN network (pre-2023)
site_order_list <- load_site_order(file_path = here("data", "derived", "auto_qaqc_files", "site_order", "site_order_pre2023.yml"))

final_flags <- intrasensor_flags_list %>%
  purrr::map(~network_check(df = ., intrasensor_flags_arg = intrasensor_flags_list, site_order_arg = site_order_list)) %>%
  rbindlist(fill = TRUE) %>%
  tidy_flag_column() %>%
  split(f = list(.$site, .$parameter), sep = "_") %>%
  purrr::map(~add_suspect_flag(.)) %>%
  data.table::rbindlist(fill = TRUE)
```

## 9) Final post-processing & save outputs

We remove isolated one-off “suspect data” points and write per-site/parameter CSVs for the reporting cycle.

```{r}
v_final_flags <- final_flags%>%
  dplyr::mutate(auto_flag = ifelse(is.na(auto_flag), NA,
                                   ifelse(auto_flag == "suspect data" & is.na(lag(auto_flag, 1)) & is.na(lead(auto_flag, 1)), NA, auto_flag))) %>%
  dplyr::select(c("DT_round", "DT_join", "site", "parameter", "mean", "units", "n_obs", "spread", "auto_flag", "mal_flag", "sonde_moved","sonde_employed", "season", "last_site_visit")) %>%
  dplyr::mutate(auto_flag = ifelse(is.na(auto_flag), NA, ifelse(auto_flag == "", NA, auto_flag))) %>%
  mutate(year = year(DT_round)) %>%
  split(f = list(.$site, .$parameter, .$year), sep = "_") %>%
  keep(~nrow(.) > 0)
```

## 10) Visualize flagged data for sanity check

```{r}

 final_plot_data <- v_final_flags%>% 
  bind_rows()%>%
  split(f = list(.$site, .$year), sep = "_")%>%
  keep(~nrow(.) > 0)
  
  plotz <- map(names(final_plot_data), ~ {
  ggplot(data = final_plot_data[[.x]]) +
    geom_point(aes(x = DT_round, y = mean, color = auto_flag), na.rm = TRUE) +
    facet_wrap(~parameter, scales = "free_y") +
    #remove legend for auto_flag
    theme(legend.key = element_blank(),
          legend.title = element_blank(),
          legend.position = "none") +
    labs(
      title = .x,
      x = "Date/Time",
      y = "Mean Value"
    )
})
  
  walk(plotz, print)
```


## 11) Save processed data

  ### Save to flagged folder

```{r}
# Save files
iwalk(v_final_flags, ~ {
  site_name <- str_split(.y, "_")[[1]][1]  # Extract just the site name (first part)
  parm <- str_split(.y, "_")[[1]][2]
  year <- str_split(.y, "_")[[1]][3]
  dir_path <- here("data", "raw", "sensor", "manual_data_verification", paste0(year, "_cycle"), "hydro_vu_pull", "flagged")
  if(!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }  
  filename <- here(dir_path, paste0(tolower(site_name), "_", parm, ".parquet"))
  .x %>% select(-year) %>% write_parquet(filename)
  cat("Saved:", filename, "\n")
})
```

### Save data into Shiny Ver Tool formatting

```{r}

data_to_ver <- v_final_flags%>%
  bind_rows()%>%
  split(f = list(.$year), sep = "_")%>%
  keep(~nrow(.) > 0)

iwalk(data_to_ver, \(x, idx) {
  message("Processing year: ", idx)
  base_dir = here("data", "raw", "sensor", "manual_data_verification", paste0(idx, "_cycle"), "in_progress")
  all_path = here(base_dir, "all_data_directory")
  pre_verification_path = here(base_dir, "pre_verification_directory")
  intermediary_path = here(base_dir, "intermediary_directory")
  verified_path = here(base_dir, "verified_directory")
  raw_data_path = here(base_dir, "raw_data")
  meta_folder_path = here(base_dir, "meta")
  
  
  #check/ create all necessary folders
  if(!dir.exists(all_path)){
    dir.create(all_path)
  }
  #check to see if data folder has pre_verification_directory
  if(!dir.exists(pre_verification_path)){
    dir.create(pre_verification_path)
  }
  #check to see if data folder has intermediary_directory
  if(!dir.exists(intermediary_path)){
    dir.create(intermediary_path)
  }
  
  #check to see if data folder has verified_directory
  if(!dir.exists(verified_path)){
    dir.create(verified_path)
  }
  #create raw data folder
  if(!dir.exists(raw_data_path)){
    dir.create(raw_data_path)
  }
  if(!dir.exists(meta_folder_path)){
    dir.create(meta_folder_path)
    #save the site order file and seasonal threshold file to the meta folder for use in the app
    file.copy(from = here("data", "derived", "auto_qaqc_files", "site_order", "site_order_pre2023.yml"), to = meta_folder_path)
    file.copy(from = here("data","derived","auto_qaqc_files", "thresholds","seasonal_thresholds20260304-T175010Z.csv"), to = meta_folder_path)
    file.rename(from = here(meta_folder_path, "seasonal_thresholds20260304-T175010Z.csv"), to = here(meta_folder_path, "seasonal_thresholds.csv"))
    #copy over the available flags file to the meta folder for use in the app
    file.copy(from = here("data","raw","sensor", "manual_data_verification","2025_cycle", "in_progress", "meta",   "available_flags.csv"), to = meta_folder_path)
    #copy over the parameter auto selections
    file.copy(from = here("data","raw","sensor", "manual_data_verification","2025_cycle", "in_progress", "meta",   "parameter_autoselections.csv"), to = meta_folder_path)
  }
  
  all_data_parsed <- x%>%
    mutate(DT_round = with_tz(DT_round, tzone = "MST"))%>%
       mutate(year = year(DT_round))%>%
    #double check we are only saving one year
    filter(year == idx)
  
  #Taking all_data_parsed and formatting to save into all and pre directories
  site_params <- expand_grid(site = unique(all_data_parsed$site), parameter = unique(all_data_parsed$parameter))
  #create a list were each is named after the site and parameter combination
  flagged_list <- pmap(site_params, ~ all_data_parsed %>% filter(site == ..1, parameter == ..2))
  rm(all_data_parsed) #save some memory here
  names(flagged_list) <- paste(site_params$site, site_params$parameter, sep = "-")
  #remove empty lists
  flagged_list <- keep(flagged_list, ~nrow(.x) > 0)
  #add additional columns needed
  pre_processed_data <- map(.x = flagged_list, \(x)
                            x %>%
                              mutate(
                                mean_verified = mean, # if verification status pass or flag, this is mean. set default to mean so that it can be used in the app
                                is_verified = FALSE, # whether or not the data has been verified
                                verification_status = NA, # can only be pass/fail/skip
                                year = year(DT_round),
                                week = week(DT_round),
                                weekday = lubridate::wday(DT_round, week_start = 7),
                                y_w = paste(year, "-", week),
                                day = yday(DT_round),
                                y_d = paste(year, "-", day),
                                #based on Colorado Hydrology, should be removed for DS version
                                season = case_when(
                                  month(DT_round) %in% c(12, 1, 2, 3, 4) ~ "winter_baseflow",
                                  month(DT_round) %in% c(5, 6) ~ "snowmelt",
                                  month(DT_round) %in% c(7, 8, 9) ~ "monsoon",
                                  month(DT_round) %in% c(10, 11) ~ "fall_baseflow"
                                ),
                                #auto_flag = flag,
                                #merge auto_flag and mal_flag by ; into flag. don't include NAs
                                flag = case_when(
                                  #both are flagged then combine by ;
                                  !is.na(auto_flag)&!is.na(mal_flag) ~ paste(auto_flag, mal_flag, sep = ";"),
                                  #only auto_flag is flagged, only use auto_flag
                                  !is.na(auto_flag)&is.na(mal_flag) ~ auto_flag,
                                  #only mal_flag is flagged, only use mal_flag
                                  !is.na(mal_flag)&is.na(auto_flag) ~ mal_flag
                                ),
                                user_flag = flag, # Keep Auto flags in flag column, user added flags/alterations live in user flag
                                brush_omit = FALSE, # User Brush Omit instances, default to FALSE
                                user = NA, #User initials
                                final_status = NA, # Final status of the data point after weekly decision
                                week_decision = NA,  #store weekly decision
                                is_finalized = FALSE) #Only TRUE if user submits final decision button
  )
  
  #Double check TZs, grab all temperature data and plot monthly averages by hour for each site
  # We should expect to see a diurnal pattern in the temperature data, with peaks around ~ 15:00 (3pm) and troughs around ~ 5:00/7:00 (5-7am). If we don't see this pattern, it could be an indication that the timestamps are not correct or that there is some other issue with the data.
  # pre_processed_data%>%
  #   keep(~ unique(.x$parameter) %in% c("Temperature"))%>%
  #   bind_rows()%>%
  #   mutate(hour = hour(DT_round))%>%
  #   group_by(site, parameter, month = month(DT_round), hour)%>%
  #   summarize(mean = mean(mean, na.rm = TRUE))%>%
  #   ggplot(aes(x = hour, y = mean, color = site))+
  #   geom_line()+
  #   facet_wrap(~month)
  
  rm(flagged_list) #save some memory here
  
  iwalk(pre_processed_data, \(x, idx) {
    timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
    data_hash <- digest::digest(x)
    new_filename <- glue::glue("{idx}_{timestamp}_{data_hash}.parquet")
    write_parquet(x, here(all_path, new_filename))  # Ensure you use new_filename, not idx
  })
  
  R.utils::copyDirectory(all_path, pre_verification_path)
  message("Files saved to all and pre directory")
})

```


# Future Work

## Drift Correction post-ver

```{r}
#Create a post_verification_directory for each cycle
years <- c(2019, 2020, 2021)

walk(years, ~ {
  dir_path <- here("data", "raw", "sensor", "manual_data_verification", paste0(.x, "_cycle"), "post_verification")
  if(!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
})

```


## Prep for HydroShare Upload

Copied from `Process_2025.rmd` but should have the same format


We will be creating two datasets here: 
- A "raw" dataset that includes all the original data with flags and all "versions" of the data: no manual corrections applied, calibration correction, post verification, and post drift correction. This is for posterity and to have a record of what the original data looked like before and after manual verification/corrections.
- A "cleaned" dataset that includes only the final value (post drift correction if it exists, post verification if not) and a column indicating what modifications were made to the data (manual exclusion, calibration correction, drift correction, etc). This is the dataset that will be uploaded to HydroShare and used for analysis.

### Prep Data

```{r}

years <- c(2019, 2020, 2021)%>%paste0("_cycle")

hydro_share_datasets <- map(years, function(year_cycle){
  
  
drift_corrected_files <- list.files(path = here("data/raw/sensor/manual_data_verification/", year_cycle,"/post_verification/"), pattern = "*.parquet", full.names = TRUE)

drift_corrected_data <- map_dfr(drift_corrected_files, 
                                ~read_parquet(.x)%>% 
                                  ross.wq.tools::add_column_if_not_exists(., "pre_post_source")%>%
                                  mutate(pre_post_source = as.character(pre_post_source)))%>%
  select(DT_round, DT_join, site, parameter,units, 
         value_raw = mean,  
         value_post_ver = mean_analysis, 
         value_post_drift_corr = mean_drift_trans,   
         auto_qc_flag = auto_flag, mal_flag, user_qc_flag = user_flag, qc_status = final_status)%>%
  split(f = list(.$site, .$parameter), sep = "-")%>%
  keep(~ nrow(.x) > 0) #remove any empty datasets

non_drift_corrected_files <- list.files(path = here("data/raw/sensor/manual_data_verification/", year_cycle, "in_progress/verified_directory/"), pattern = "*.parquet", full.names = TRUE)

non_drift_corrected_data <- map_dfr(non_drift_corrected_files, ~read_parquet(.))%>%
  mutate( mean_analysis = ifelse(verification_status %in% c("FLAGGED", "PASS"), mean, NA))%>%
  select(DT_round, DT_join, site, parameter,units,
         value_raw = mean, 
         value_post_ver = mean_analysis,
         auto_qc_flag = auto_flag, 
         mal_flag, 
         user_qc_flag = user_flag,
         qc_status = final_status)%>%
  split(f = list(.$site, .$parameter), sep = "-")

#only keep non drift corrected site parameters if they are not in the drift corrected data (i.e. if they haven't been drift corrected yet)
all_datasets <- c(non_drift_corrected_data[!names(non_drift_corrected_data) %in% names(drift_corrected_data)], drift_corrected_data)%>%
  keep(~ nrow(.x) > 0) #remove any empty datasets


#units lookup table

units_table <- all_datasets%>%
  bind_rows()%>%
  select(parameter, units)%>%
  distinct()%>%
  filter(!is.na(parameter) & !is.na(units))

raw_dataset <- all_datasets%>%
  bind_rows()%>%
  # if value_post_drift corr is NA, then fill with value_post_ver (this means that the data was verified but not drift corrected, so the post verification value is the most "final" version of the data we have for those instances) This is for parameters other than FDOM and Turbidity
  mutate(value_post_drift_corr = if_else(is.na(value_post_drift_corr), value_post_ver, value_post_drift_corr),
         #round all values to 3 decimal places for consistency with HydroShare upload
         value_raw = round(value_raw, 3),
         value_post_ver = round(value_post_ver, 3),
         value_post_drift_corr = round(value_post_drift_corr, 3),
         #Create modifications column to indicate what modifications were made to the data for easier filtering and analysis downstream
         modifications = case_when( is.na(value_post_ver) & !is.na(value_raw) ~ "manual data exclusion",
                                    value_post_ver != value_post_drift_corr ~ "drift correction",
                                    TRUE ~ NA), 
         #convert to UTC
         DateTime_UTC = with_tz(DT_round, tzone = "UTC"))%>%
  select(-units)%>%
  left_join(units_table, by = "parameter")%>%
  select(DateTime_UTC,
         site, parameter, units, 
         #Actual data
         value_raw, 
         value_post_ver,
         value_post_drift_corr, 
         #QC Info
         auto_qc_flag, mal_flag, user_qc_flag, qc_status, modifications)%>%
  distinct(.keep_all = TRUE)
    
final_dataset <- raw_dataset %>%
  select(DateTime_UTC, site, parameter, units, value_raw, value_final = value_post_drift_corr,user_qc_flag, modifications)%>%
  distinct(.keep_all = TRUE)

return(list(raw_dataset = raw_dataset, final_dataset = final_dataset))
})%>%
  set_names(years)
```

#### Plot to Double Check 

```{r}
# 
final_dataset <- map(hydro_share_datasets, "final_dataset")%>%
  bind_rows()%>%
  filter(!is.na(value_final))

plots <- map(unique(final_dataset$site), function(site_name) {
  ggplot(final_dataset %>% filter(site == site_name), aes(x = DateTime_UTC, y = value_final, color = modifications)) +
    geom_point(aes(y = value_raw), alpha = 0.5) + # add raw data points for comparison
    geom_point() +
    facet_wrap(~parameter, scales = "free_y") +
    labs(title = site_name,
         x = "Date",
         y = "Value",
         color = "QC Status")
})%>%
  set_names(unique(final_dataset$site))
walk(plots, print)


#plot specific site and parameter
final_dataset%>%
  filter(site == "archery", parameter == "Specific Conductivity")%>%
  ggplot(aes(x = DateTime_UTC, y = value_final, color = modifications)) +
  geom_point(aes(y = value_raw), alpha = 0.5) + # add raw data points for comparison
  geom_point() +
  labs(title = "Archery - Turbidity",
       x = "Date",
       y = "Value",
       color = "QC Status")
```


### Save Data

```{r}

for(year in c("2019", "2020", "2021")){
  year_cycle = paste0(year, "_cycle")
  
  final_dataset <- hydro_share_datasets[[year_cycle]][["final_dataset"]]
  raw_dataset <- hydro_share_datasets[[year_cycle]][["raw_dataset"]]

# 
#check if the folder exists, if not create it
if (!dir.exists(here( "data/collated/sensor/hydroshare_pub", year))) {
  dir.create(here( "data/collated/sensor/hydroshare_pub", year), recursive = TRUE)
}

cleaned_file_name <- here( "data/collated/sensor/hydroshare_pub", year, paste0("CSU_ROSS_", year,"_cleaned.csv"))

raw_file_name <- here( "data/collated/sensor/hydroshare_pub", year, paste0("CSU_ROSS_", year,"_raw.csv"))

write_csv(final_dataset, cleaned_file_name)

write_csv(raw_dataset, raw_file_name)
# Repeat for parquet files
cleaned_file_name_parquet <- here( "data/collated/sensor/hydroshare_pub", year, paste0("CSU_ROSS_", year,"_cleaned.parquet"))

raw_file_name_parquet <- here( "data/collated/sensor/hydroshare_pub", year, paste0("CSU_ROSS_", year,"_raw.parquet"))

write_parquet(final_dataset, cleaned_file_name_parquet)

write_parquet(raw_dataset, raw_file_name_parquet)
}
```

## Field Notes

```{r}

old_field_notes <- readxl::read_excel(here("data","raw","field_notes","sensor_field_notes.xlsx")) %>% 
  mutate(DT = (paste0(date, " ", start_time_mst))) %>%
  mutate(DT = ymd_hm(DT) + hours(7)) %>%
  arrange(DT) %>%
  mutate(DT_round = round_date(DT, "15 minutes")) %>%
  mutate(DT_round = with_tz(DT_round, tzone = "UTC"),
         last_site_visit = with_tz(DT_round, tzone = "UTC"),
         DT_join = as.character(DT_round),
         sonde_moved = NA,
         sonde_employed = case_when(is.na(sensor_deployed) & is.na(sensor_pulled) ~ NA,
                                    sensor_deployed == "x" ~ 0, 
                                    sensor_pulled == "x" ~ 1), 
         site = ifelse(site == "rist", "bellvue", site)) %>%
  ross.wq.tools::fix_site_names()%>%
  #add extra cols
  mutate(rdo_pre_clean = NA, 
         rdo_post_clean = NA,
         rdo_post_cal = NA,
         cond_pre_clean = NA,
         cond_post_clean = NA,
         cond_post_cal = NA,
         ph_pre_clean = NA,
         ph_post_clean = NA,
         ph_post_cal = NA,
         orp_pre_clean = NA,
         orp_post_clean = NA,
         orp_post_cal = NA,
         turb_pre_clean = NA,
         turb_post_clean = NA,
         turb_post_cal = NA,
         fdom_pre_clean = NA,
         fdom_post_clean = NA,
         fdom_post_cal = NA,
         chla_pre_clean = NA,
         chla_post_clean = NA,
         chla_post_cal = NA)%>%
  select(-sensors_cleaned, -wiper_working, -cal_report_collected)%>%
  filter(DT_round <= ymd("2021-12-31"))

formatted_notes <- old_field_notes%>%
  filter(!(site %in% c("springcreek", "boxcreek", "mountaincampus")))%>%
  rename(DateTime_UTC = DT_round) %>%
  select(-any_of(c("photos_downloaded","DT_join", "cal_report_collected", "sensor_swapped_notes", "end_dt", "last_site_visit", "date","sensor_deployed", "sensor_pulled")))%>%
  select(any_of(c("site",  "crew","DateTime_UTC", "sonde_employed", "sensors_cleaned", "wiper_working",
                  "rdo_cap_condition", "rdo_cap_replaced", "ph_junction_replaced", "cals_performed",
                  "sonde_moved", "sensor malfunction", "visit_type", "start_DT", "visit_comments", 
                  "rdo_pre_clean",   "rdo_post_clean", "rdo_post_cal",
                  "cond_pre_clean", "cond_post_clean", "cond_post_cal",
                  "ph_pre_clean", "ph_post_clean", "ph_post_cal",
                  "orp_pre_clean", "orp_post_clean", "orp_post_cal",
                  "turb_pre_clean", "turb_post_clean", "turb_post_cal",
                  "fdom_pre_clean", "fdom_post_clean", "fdom_post_cal",
                  "chla_pre_clean", "chla_post_clean", "chla_post_cal")))%>%
  #Clean up crew to just initials and remove any extra text
  mutate(crew = crew %>%
    str_replace_all("Other \\(please specify\\)", "Other") %>%
    str_replace_all("([A-Z])[a-z]+", "\\1") %>%
    str_replace_all("(?<=[A-Z])\\s+(?=[A-Z])", "")
  )%>%
  fix_site_names()

#Only 2021 had field notes
  
  field_notes_file_name <- here( "data/collated/sensor/hydroshare_pub/2021/CSU_ROSS_2021_field_notes.csv")
  
  write_csv(formatted_notes, field_notes_file_name)

```



## Location Metadata

Here we will load and double check the location metadata file created by Sam Struthers and verified by Katie Willi

```{r}
location_meta <- read_csv("data/collated/sensor/hydroshare_pub/CSU_ROSS_sonde_location_metadata.csv")

# location_meta <- read_csv("data/collated/sensor/hydroshare_pub/CSU_ROSS_sonde_location_metadata.csv",
#                            show_col_types = FALSE)%>%
#   mutate(data_start_date = as.POSIXct(data_start_date, tz = "UTC", format = "%m/%d/%y"),
#          data_end_date =  if_else( is.na(data_end_date), 
#                                    ymd("2025-11-17"),
#                                    as.POSIXct(data_end_date, tz = "UTC", format = "%m/%d/%y")))%>%
# 
#   mutate(min_year = year(data_start_date),
#          max_year = year(data_end_date), 
#          year_range = paste0(min_year, "-", max_year))
# #update as needed
# #write_csv(location_meta, here("data/collated/sensor/hydroshare_pub/CSU_ROSS_sonde_location_metadata.csv"))
# 
# tmap_mode("view")
# tm_basemap("OpenStreetMap") +
#   tm_shape(location_meta %>% st_as_sf(coords = c("longitude", "latitude"), crs = 4326)) +
#   #color is SiteID and shape is year_range
#   tm_symbols(col = "SiteID", shape = "year_range", size = 0.5, border.col = "black")

```



