---
title: "burial_flag_method_comparison_jpd_09-10-26"
output: html_document
---

# Motivation

Our current DO burial detection is flagging good data as burials. The DO
interference heuristics we came up with are too coarse to tell the
difference between data that is just unusual and data that actually
indicates a sonde burial. The heuristics check for two patterns in the DO
signal and tag either one as "DO interference". The first is spikes in the
data, which serve as an analog for increased variability in the signal. This type of signal may be indicative of a partial burial as sediment may fill in near the sensor and periodically flush out.The
second is "unrealistic" lows in the data, a depressed signal that is either
not plausible or only possible as a reading because the DO sensor is buried
and has limited access to moving oxygenated water. By combining these two
signals we had previously been able to detect burials with reasonable
precision and recall.

The water year regimes we built these tools in had rarely if ever seen DO
levels lower than 5 mg/L that were not burials, but 2026 has been an
extremely anomalous year in terms of precipitation and elevated water temperatures. As a result we have seen
realistic values lower than 5 mg/L that trigger our DO burial heuristics and
flag large swaths of correct data. Given that these functions are used in
live environments with users who rely on the correctness of our flagging
methods, we want to correct this behavior. In particular, we want to reduce "false positives" with this flag (ie good data being flagged as a burial)

# Data Loading and Environment Initialization

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(arrow)
library(ross.wq.tools)
library(plotly)
library(here)
library(slider)
```

To explore new flagging functions, we will use data collected in 2023 through 2026. For 2023 to 2025 we use the published HydroShare
`raw` files, which carry `user_qc_flag` (the human verification outcome, our
ground truth) alongside `auto_qc_flag` (what the current auto-QAQC produced).
2026 is not verified yet, so we pull it from the in-progress
manual-verification directory. It has no `user_qc_flag`, so we set that column
to NA and line the rest of the schema up with the older years. Column
definitions are in `data/collated/sensor/hydroshare_pub/README.md` and the flag
definitions in `CSU_ROSS_user_qc_flag_metadata.csv`. Earlier years are left out
because they carry no site-visit information, which the updated method needs.

```{r load-do}
# 2023-2025: published HydroShare raw files
do_hydro <- list.files(
  here("data", "collated", "sensor", "hydroshare_pub"),
  pattern = "CSU_ROSS_[0-9]{4}_raw\\.parquet$",
  full.names = TRUE, recursive = TRUE
) %>%
  keep(~ as.integer(str_extract(basename(.x), "[0-9]{4}")) %in% 2023:2025) %>%
  map(~ read_parquet(.x) %>% filter(parameter == "DO")) %>%
  bind_rows() %>%
  transmute(
    site, parameter,
    DT_round    = DateTime_UTC,
    mean        = value_raw,
    auto_flag   = auto_qc_flag,
    mal_flag, user_qc_flag,
    # ground truth: technician-verified burial. Catches "sonde burial" and
    # "reported sonde burial" (field-note derived), not "Possible burial",
    # which is the algorithm's own output.
    user_burial = str_detect(replace_na(user_qc_flag, ""), "sonde burial")
  )

# 2026: in-progress manual-verification pull, one file per site-parameter
raw_2026 <- list.files(
  here("data", "raw", "sensor", "manual_data_verification", "2026_cycle",
       "hydro_vu_pull", "flagged_data"),
  pattern = "-DO\\.parquet$", full.names = TRUE
) %>%
  map(read_parquet) %>%
  bind_rows()

do_2026 <- raw_2026 %>%
  transmute(
    site, parameter, DT_round, mean, auto_flag, mal_flag,
    user_qc_flag = NA_character_,
    # no human verification yet, so fall back to the malfunction flag
    user_burial  = str_detect(replace_na(mal_flag, ""), "burial")
  )

# 2026 also carries last_site_visit, which we use below to mark visits
visits_2026 <- raw_2026 %>% distinct(site, visit_DT = last_site_visit)

do_data <- bind_rows(do_hydro, do_2026) %>%
  ross.wq.tools::fix_site_names() %>%
  mutate(year = lubridate::year(DT_round)) %>%
  arrange(site, DT_round)
```

Now prep the data: pull in the field notes for the visit timestamps, mark each
site visit in the series, then split by year and by site so the methods can run
over one site-year at a time.

```{r prep-do}
# field notes give visit times that were not always written into a flag column
field_note_visits <- c("2023", "2024", "2025") %>%
  map_dfr(~ read_csv(
    here("data", "collated", "sensor", "hydroshare_pub", .x,
         sprintf("CSU_ROSS_sensor_field_notes_%s.csv", .x)),
    show_col_types = FALSE
  ) %>% transmute(site, visit_DT = DateTime_UTC))

visits <- bind_rows(field_note_visits, visits_2026) %>%
  ross.wq.tools::fix_site_names() %>%
  filter(!is.na(visit_DT)) %>%
  distinct(site, visit_DT)

# a row is a site visit if a flag column says so, or if it is the closest
# observation to a field-note visit time. Mark only the first row of each run
# so the visit split increments once per visit.
do_prepped <- do_data %>%
  group_by(site) %>%
  group_modify(function(df, key) {
    v <- visits$visit_DT[visits$site == key$site]
    nearest <- map_int(v, function(t) {
      #Find the nearest site visit within 180 minutes (3 hours)
#Find the nearest site visit within 180 minutes (3 hours)
      d <- as.numeric(abs(difftime(df$DT_round, t, units = "mins")))
      if (min(d) > 180) NA_integer_ else which.min(d)
    })
    df %>% mutate(
      flag_txt   = paste(replace_na(auto_flag, ""), replace_na(user_qc_flag, "")),
      hit        = str_detect(flag_txt, "site visit(?! window)") |
                     row_number() %in% nearest,
      flag       = if_else(hit & !lag(hit, default = FALSE),
                           "site visit", NA_character_),
      site_visit = !is.na(flag)
    ) %>% select(-flag_txt, -hit)
  }) %>%
  ungroup()

# do_split[["2024"]][["elc"]] -> one site-year DO frame
do_split <- do_prepped %>%
  filter(year %in% 2023:2026) %>%
  split(f = .$year) %>%
  map(~ split(.x, f = .x$site))
```

# Method comparison

Here we compare the original method with an updated version of the method
that Sam came up with. I have slightly tweaked the method to reduce
the function's false positive rate (reduce the amount of flags we put on 
data that should not be flagged).

## Original Method

The original method is `find_do_noise()` in `ross.wq.tools`. It works in two
steps. First it tags each point as "DO interference" if the point sits at
least 0.5 mg/L below both of its neighbors (a spike), or if DO is at or below
5 mg/L. Then it tags "Possible burial" wherever a 96-observation window (about
24 hours) holds at least 24 interference points. That window check runs both
right-aligned and center-aligned, and a point is flagged if either alignment
trips.

The window is counted in observations, not in time, and the series is never
split at site visits. So a data gap shrinks the real time span the window
covers, and a window can average across a cleaning event.

```{r original-method}
# Original method: find_do_noise() from ross.wq.tools, written out step by step
# instead of called. Two steps, run per site-year so a window never crosses a
# site or a year boundary:
#   1. point-level "DO interference": the point sits >= 0.5 mg/L below both of
#      its neighbors (a spike), or DO is at or below 5 mg/L.
#   2. "Possible burial": a 96-observation (~24 h) window holds >= 24
#      interference points, checked both right- and center-aligned.
# No site-visit split. The window is counted in observations, not in time.
# Every year is flagged the same way. 2023 is the year with the most confirmed
# real burials; 2026 is the anomalous low-DO year from the motivation.

DO_old <- do_split %>%
  map(~ map(.x, function(df) {

    df <- df %>%
      arrange(DT_round) %>%
      generate_summary_statistics()          # adds back1 / front1 (the neighbors)

    df %>%
      mutate(
        # --- step 1: point-level interference -------------------------------
        # downward spike: this point is >= 0.5 mg/L below the point before it
        # and the point after it
        old_spike = (back1 - mean >= 0.5) & (front1 - mean >= 0.5),
        # unrealistic low: DO at or below 5 mg/L
        old_low   = mean <= 5,
        old_interference = replace_na(old_spike | old_low, FALSE),

        # --- step 2: persistence over ~24 h --------------------------------
        # count interference points in a rolling 96-observation window and flag
        # the window if it holds at least 24. Run it right-aligned (window ends
        # on the point) and center-aligned (point in the middle).
        .bin = as.integer(old_interference),
        .win_right  = data.table::frollapply(.bin, N = 96,
                        FUN = function(x) sum(x) >= 24, align = "right",  fill = NA),
        .win_center = data.table::frollapply(.bin, N = 96,
                        FUN = function(x) sum(x) >= 24, align = "center", fill = NA),
        old_burial = replace_na(.win_right == 1 | .win_center == 1, FALSE),

        # combined flag text, same format the package writes
        flag_old = case_when(
          old_interference & old_burial ~ "DO interference;Possible burial",
          old_burial                    ~ "Possible burial",
          old_interference              ~ "DO interference",
          TRUE                          ~ NA_character_
        )
      ) %>%
      select(-.bin, -.win_right, -.win_center)
  })) %>%
  list_flatten() %>%
  list_rbind()
```

## Updated Method

The updated method is Sam's `find_do_noise_new()` with one tweak from me. It
first splits the series at every site visit so no window ever averages across
a cleaning. Within each split it computes three windows:

- `low_right`: mean DO below 5 mg/L over the trailing 24 hours
- `low_center`: mean DO below 5 mg/L over a centered +/- 12 hour window
- `static_right`: DO range (max minus min) under 1 mg/L over the trailing 24
  hours, meaning the diel cycle is gone. Needs at least 10 points in the
  window.

Sam's flag reduces to `low_right | low_center`.

My tweak is on `low_center`. At the end of a split the centered window loses
its forward half, so a point sitting in the normal overnight low can end up
with a trailing-only window that averages below 5 mg/L. That flags the last
few points before a site visit as burial when nothing is wrong. With the
tweak, `low_center` can flag on its own only when its +/- 12 hour window
actually holds data on both sides (`center_well_covered`, at least 80% of the 720 minute half-span,
tweak, `low_center` can flag on its own only when its +/- 12 hour window
actually holds data on both sides (`center_well_covered`, at least 80% of the 720 minute half-span,
one of the other two signals. The flag becomes
`low_right | (static_right & low_center) | (low_center & center_well_covered)`.

```{r updated-method}
# Updated method: Sam's find_do_noise_new() plus my low_center coverage gate,
# written out step by step. Run per site-year, and within each, split the series
# at every site visit so no window averages across a cleaning.
#   low_right    : trailing 24 h mean DO < 5 mg/L
#   low_center   : centered +/- 12 h mean DO < 5 mg/L
#   static_right : trailing 24 h DO range < 1 mg/L (diel cycle gone), >= 10 pts
#   center_well_covered : the +/- 12 h window holds >= 80% of its half-span of
#     data on BOTH sides -- my gate on a lone low_center
# Two burial columns are kept so the tweak's effect can be measured:
#   new_burial_sam = low_right | low_center                       (Sam's)
#   new_burial     = low_right | (static_right & low_center)
#                    | (low_center & center_well_covered)         (with my tweak)

too_low       <- function(x) mean(x, na.rm = TRUE) < 5
no_diel_cycle <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) < 10) return(NA)          # not enough points to judge the range
  (max(x) - min(x)) < 1.0
}

DO_new <- do_split %>%
  map(~ map(.x, function(df) {

    half_span_secs <- 720 * 60             # +/- 12 h, in seconds
    min_cover      <- 0.8                  # 80% of each half-span, measured in time

    df %>%
      arrange(DT_round) %>%
      generate_summary_statistics() %>%
      # split at each site visit
      mutate(visit_group = cumsum(str_detect(replace_na(flag, ""), "site visit"))) %>%
      group_by(visit_group) %>%
      mutate(
        low_right    = slide_index_dbl(mean, DT_round, too_low,
                         .before = minutes(1440)),
        low_center   = slide_index_dbl(mean, DT_round, too_low,
                         .before = minutes(720), .after = minutes(720)),
        static_right = slide_index_dbl(mean, DT_round, no_diel_cycle,
                         .before = minutes(1440)),
        # time actually spanned by the centered window on each side of the point
        span_before  = slide_index_dbl(as.numeric(DT_round), DT_round,
                         ~ max(.x) - min(.x), .before = minutes(720)),
        span_after   = slide_index_dbl(as.numeric(DT_round), DT_round,
                         ~ max(.x) - min(.x), .after = minutes(720))
      ) %>%
      ungroup() %>%
      mutate(
        lr = replace_na(low_right    == 1, FALSE),
        lc = replace_na(low_center   == 1, FALSE),
        sr = replace_na(static_right == 1, FALSE),
        center_well_covered = replace_na(
          span_before >= min_cover * half_span_secs &
          span_after  >= min_cover * half_span_secs, FALSE),

        new_burial_sam = lr | lc,
        new_burial     = lr | (sr & lc) | (lc & center_well_covered),
        flag_new = if_else(new_burial, "Possible burial", NA_character_)
      ) %>%
      select(-lr, -lc, -sr)
  })) %>%
  list_flatten() %>%
  list_rbind()
```


# Method Comparisons

These two methods go after the same problem in very different ways. The point
I want to make here is that there is no silver bullet. A burial event shows up
in many different forms in the DO signal, and both methods try to capture that
variability with heuristics built from physical rules. The solution space is
huge. Even these two methods have many permutations I did not test, and many
more tweaks could be added to make them "better". 

The table below lays out the physical reasoning behind each rule and which
method uses it.

| Rule | Physical reasoning | Original | Updated |
|------|--------------------|----------|---------|
| Point sits well below both neighbors (spike) | DO cannot change that fast and bounce back; a lone drop is a sensor artifact, not water chemistry | yes | no |
| DO below 5 mg/L | a buried sensor loses contact with moving, oxygenated water and reads too low | yes (per point) | yes (window mean) |
| Low DO persists for ~24 h | a real burial lasts; transient noise does not | yes (count of interference points) | yes (window mean) |
| Diel cycle is gone (flat signal) | burial removes the daily photosynthesis and respiration swing in DO | no | yes (`static_right`) |
| Split the series at each site visit | a cleaning removes sediment from the sensor, so data on either side should not be pooled | no | yes |
| Centered window, not just trailing | a centered window catches the onset of a burial about 12 h earlier | no | yes (`low_center`) |
| Ignore a centered window missing half its data | a half-empty window near a visit or gap should not flag on its own | no | yes (my tweak) |
| Windows measured in time, not observations | a data gap should not silently shrink the real time span the window covers | no | yes |

```{r join-methods}
# one row per site-timestamp, both methods side by side
DO_both <- DO_old %>%
  select(site, DT_round, year, mean, season, site_visit, user_burial,
         old_interference, old_burial, flag_old) %>%
  left_join(
    DO_new %>% select(site, DT_round, visit_group, low_right, low_center,
                      static_right, center_well_covered,
                      new_burial_sam, new_burial, flag_new),
    by = c("site", "DT_round")
  ) %>%
  # which subset of {original, updated} flagged each point
  mutate(combo = factor(
    case_when(
      !old_burial & !new_burial ~ "neither",
       old_burial & !new_burial ~ "original only",
      !old_burial &  new_burial ~ "updated only",
       old_burial &  new_burial ~ "both"
    ),
    levels = c("neither", "original only", "updated only", "both")
  )) %>%
  # is the point within 13 h of a site visit? (for the near-visit analysis)
  group_by(site, year) %>%
  mutate(near_visit = {
    vt <- sort(as.numeric(DT_round[site_visit]))
    if (length(vt) == 0) rep(FALSE, n()) else {
      t <- as.numeric(DT_round); j <- findInterval(t, vt)
      d_prev <- ifelse(j >= 1,        t - vt[pmax(j, 1)],              Inf)
      d_next <- ifelse(j < length(vt), vt[pmin(j + 1, length(vt))] - t, Inf)
      pmin(d_prev, d_next) < 13 * 3600
    }
  }) %>%
  ungroup()
```

```{r plot-fn}
# One interactive plot for a site-year: DO over time, points colored by which
# method(s) flagged "Possible burial", red dashed lines at site visits, orange
# shading over verified-burial stretches (user_qc_flag, or mal_flag for 2026).
plot_burial <- function(site_name, yr, data = DO_both) {

  d <- data %>% filter(site == site_name, year == yr) %>% arrange(DT_round)
  if (nrow(d) == 0) stop("no data for ", site_name, " ", yr)

  # verified-burial stretches -> plotly rectangles (contiguous runs, gap > 20 min
  # starts a new one). ggplotly drops geom_rect(ymin = -Inf), so add them as
  # native shapes with numeric (epoch-second) x coords.
  gt <- d %>% filter(user_burial)
  shapes <- if (nrow(gt) == 0) list() else {
    gt %>%
      mutate(run = cumsum(c(TRUE, diff(as.numeric(DT_round)) > 1200))) %>%
      group_by(run) %>%
      summarise(x0 = min(DT_round), x1 = max(DT_round), .groups = "drop") %>%
      purrr::pmap(function(run, x0, x1) list(
        type = "rect", xref = "x", yref = "paper",
        x0 = as.numeric(x0), x1 = as.numeric(x1), y0 = 0, y1 = 1,
        fillcolor = "orange", opacity = 0.15, line = list(width = 0)))
  }

  combo_cols <- c("neither" = "grey80", "original only" = "#e41a1c",
                  "updated only" = "#377eb8", "both" = "#984ea3")

  p <- ggplot(d, aes(DT_round, mean, color = combo,
                     text = paste0(format(DT_round), "<br>DO: ", round(mean, 2),
                                   "<br>", combo))) +
    geom_point(size = 1) +
    geom_vline(data = filter(d, site_visit), aes(xintercept = DT_round),
               color = "red", linetype = "dashed", inherit.aes = FALSE) +
    scale_color_manual(values = combo_cols, drop = FALSE) +
    labs(title = paste(site_name, yr), x = NULL, y = "DO (mg/L)", color = NULL) +
    theme_minimal()

  gp <- plotly::ggplotly(p, tooltip = "text")
  gp$x$layout$shapes <- shapes
  gp
}
```

```{r}
plot_burial("elc", 2023)
```

# Newer Method

Here I want to show a method I found while working on this problem. It is not
something I plan to implement soon, but it is a strong approach and one we
should consider going forward. It is the third method here, so
`find_do_noise_v3()`.

## The van Brakel z-score filter

The base of the method is van Brakel's z-score filter. It keeps a running
mean and standard deviation over a trailing window of points that have so far
looked clean. A new point counts as a spike when it lands more than k standard
deviations from that running mean, where k is a multiplier (we use 4). A point
flagged as a spike does not update the running mean and standard deviation; it
is held flat instead, so a spike cannot pull the band toward itself. A point
above the band is a positive spike, below it a negative spike.

## The Skalvik sequential filter

Skalvik et al. (2024) use van Brakel's filter as one stage of a sequential
filter. The stages run in order, and a point that fails an earlier stage is
left out of the statistics every later stage is built from:

1. Drop NA and NaN.
2. High rate of change: reject a point when the step from the previous point
   is larger than a physically plausible slope.
3. van Brakel z-score band, run on the points that passed stages 1 and 2.

Skalvik also has a static out-of-range stage and a rule that the filter should
only start once the sensor is clean, deployed, and stable.

## Our tweaks

On top of the Skalvik pipeline:

- **Time-indexed windows.** The running windows are measured in time, not in
  observation counts, so a data gap does not silently shrink the window. Run
  both right-aligned and centered.
- **`sigma_nat`, a floor on the band width.** During a flat stretch the
  standard deviation collapses toward zero and every small wiggle reads as a
  spike. `sigma_nat` is the smallest the band half-width is allowed to be. It
  is derived from the 12 hour rolling standard deviation of verified-good DO
  and set by season (snowmelt 0.40, monsoon 0.50, fall baseflow 0.70, winter
  baseflow 0.80 mg/L).
- **`band_hw_max`, a cap on the band width.** Over a 24 hour window k times the
  standard deviation tracks the diel swing and can push the band out to +/- 6
  to 8 mg/L, which lets real spikes through. The half-width is capped at 2
  mg/L, so the band is `min(max(k * sd, sigma_nat), 2)`. The lower edge is
  also floored at 0, since a negative DO band means nothing.
- **Count gate and burn-in.** Stage 3 only runs with enough valid points in
  the window and only after 6 hours past the last reset.
- **Site-visit reset.** The running mean and standard deviation, the filtered
  series, and the burn-in clock all restart at each cleaning event.
- **Drop the static out-of-range stage.** The only hard bound kept is DO below
  5 mg/L, and it is not a flag on its own. It only matters together with a
  negative spike.
- **Persistence layer for burial.** Per-point signals are -1, 0, or +1.
  "Possible burial" needs a sustained run of negative spikes that are also
  below 5 mg/L over a 24 hour window (at least 75% of the evaluated points).

An earlier draft also flagged a variance collapse, where the window standard
deviation drops below a fraction of the seasonal value. Burial shows up as
both low and high variance depending on the site, so that rule was dropped.
The method is trying to do one thing: find a run of negative spikes at a
depressed baseline.

```{r v3-params}
# sigma_nat: the smallest the spike band half-width is allowed to be (mg/L).
# Derived from the 12 h rolling SD of verified-good DO, by season.
do_sigma_nat <- c(snowmelt = 0.40, monsoon = 0.50,
                  fall_baseflow = 0.70, winter_baseflow = 0.80)
do_sigma_nat_fixed <- 0.5   # fallback when a season is missing

do_v3_defaults <- list(
  k                = 3,     # SD multiplier for the band
  sigma_nat        = do_sigma_nat,
  window_hr        = 12,    # trailing window for the running mean / SD
  burn_in_hr       = 6,     # no spike verdict until this long past a reset
  min_frac_valid   = 0.5,   # window must be at least this fraction valid points
  slope_step_mult  = 15,    # t_slope_behind_* is mg/L/min -> * 15 for a 15-min step
  y_diff_max       = 1.0,   # stage-2 fallback if the slope threshold is missing
  influence        = 0,     # van Brakel w; 0 = spikes fully excluded from stats
  band_hw_max      = 2,     # hard cap on the band half-width (mg/L)
  low_do_cut       = 5,     # a negative spike only counts toward burial below this
  burial_window_hr = 24,    # persistence window for "Possible burial"
  burial_min_frac  = 0.75   # this fraction of the window low + negative -> burial
)

#' V3 DO noise / burial detector (Skalvik pipeline).
#' @param df  one site DO frame: DT_round, mean, season, flag, parameter
#' @param seasonal_thresholds  site x season table with t_slope_behind_01/99
#'   (t_mean01/99 are read but not used by the detector itself)
find_do_noise_v3 <- function(df, seasonal_thresholds, pars = do_v3_defaults) {

  if (!("DO" %in% df$parameter)) return(df)
  visit_regex <- "(site\\s*visit|\\bsv\\b)"
  site_i <- df$site[1]

  # tolerate a minimal / partial thresholds table: a missing site or missing
  # t_slope_behind_* columns degrade stage 2 to the y_diff_max fallback.
  st <- seasonal_thresholds
  if ("parameter" %in% names(st)) st <- dplyr::filter(st, parameter == "DO")
  st <- st %>%
    dplyr::filter(site == site_i) %>%
    dplyr::distinct(season, .keep_all = TRUE) %>%
    dplyr::select(dplyr::any_of(c("season", "t_mean01", "t_mean99",
                                  "t_slope_behind_01", "t_slope_behind_99")))
  for (col in c("t_mean01", "t_mean99",
                "t_slope_behind_01", "t_slope_behind_99")) {
    if (!col %in% names(st)) st[[col]] <- NA_real_
  }

  df <- df %>%
    dplyr::select(-dplyr::any_of(c("t_mean01", "t_mean99", "t_sd_0199",
                                   "t_slope_behind_01", "t_slope_behind_99"))) %>%
    dplyr::arrange(DT_round) %>%
    dplyr::mutate(visit_group = cumsum(stringr::str_detect(
      tidyr::replace_na(flag, ""),
      stringr::regex(visit_regex, ignore_case = TRUE)))) %>%
    dplyr::left_join(st, by = "season")

  n   <- nrow(df)
  y   <- df$mean
  tt  <- as.numeric(df$DT_round)                 # epoch seconds
  grp <- df$visit_group

  # per-row sigma_nat (by season, or a scalar)
  sig_nat <- if (is.null(names(pars$sigma_nat))) {
    rep(as.numeric(pars$sigma_nat[1]), n)
  } else {
    v <- unname(pars$sigma_nat[df$season]); v[is.na(v)] <- do_sigma_nat_fixed; v
  }

  # ---- stage 1: NA ----
  na_bad <- is.na(y)

  # ---- stages 2 & 3: sequential, reset per visit group ----
  rate_bad  <- logical(n)
  signal    <- integer(n)                        # -1 / 0 / +1
  avg_f     <- rep(NA_real_, n)
  sd_win    <- rep(NA_real_, n)
  band_hw   <- rep(NA_real_, n)
  filt      <- rep(NA_real_, n)                  # filteredY: survivors only
  evaluated <- logical(n)

  win_sec  <- pars$window_hr * 3600
  burn_sec <- pars$burn_in_hr * 3600
  exp_pts  <- pars$window_hr * 4                 # 15-min cadence

  for (g in unique(grp)) {
    idx  <- which(grp == g)
    seg0 <- tt[idx[1]]
    prev <- NA_integer_                          # previous point past stage 1
    last_filt <- NA_real_
    for (i in idx) {

      if (na_bad[i]) next                        # do NOT advance prev past a reject

      # stage 2: high rate vs the previous survivor, if adjacent in time
      if (!is.na(prev) && (tt[i] - tt[prev]) <= 2 * 900) {
        ymax <- suppressWarnings(max(abs(df$t_slope_behind_01[i]),
                                     df$t_slope_behind_99[i], na.rm = TRUE)) *
          pars$slope_step_mult
        if (!is.finite(ymax)) {
          ymax <- if (is.null(pars$y_diff_max)) Inf else pars$y_diff_max
        }
        if (is.finite(ymax) && abs(y[i] - y[prev]) > ymax) rate_bad[i] <- TRUE
      }
      if (rate_bad[i]) { prev <- i; next }

      # stage 3: spike vs running stats of filteredY in the trailing window
      w <- which(!is.na(filt) & tt >= tt[i] - win_sec & tt < tt[i])
      if (length(w) >= max(3, pars$min_frac_valid * exp_pts) &&
          (tt[i] - seg0) >= burn_sec) {
        mu <- mean(filt[w]); s <- stats::sd(filt[w])
        half <- max(pars$k * s, sig_nat[i])
        if (!is.null(pars$band_hw_max)) half <- min(half, pars$band_hw_max)
        avg_f[i] <- mu; sd_win[i] <- s; band_hw[i] <- half; evaluated[i] <- TRUE
        if      (y[i] > mu + half)         signal[i] <-  1L
        else if (y[i] < max(mu - half, 0)) signal[i] <- -1L  # band floor >= 0
      }

      # update filteredY (influence 0 -> a spike is carried flat, not chased)
      if (signal[i] != 0 && !is.na(last_filt)) {
        filt[i] <- pars$influence * y[i] + (1 - pars$influence) * last_filt
      } else {
        filt[i] <- y[i]
      }
      last_filt <- filt[i]
      prev <- i
    }
  }

  # ---- burial layer: sustained run of low negative spikes, reset per group ----
  low_cut   <- if (is.null(pars$low_do_cut)) 5 else pars$low_do_cut
  below_cut <- !is.na(y) & y < low_cut
  neg_ind   <- ifelse(evaluated & signal == -1 & below_cut, 1,
               ifelse(evaluated, 0, NA_real_))

  df %>%
    dplyr::mutate(
      .signal = signal, .evaluated = evaluated, .rate_bad = rate_bad,
      .avg_f = avg_f, .sd_win = sd_win, .band_hw = band_hw,
      .band_lo = pmax(avg_f - band_hw, 0), .band_hi = avg_f + band_hw,
      .neg_ind = neg_ind
    ) %>%
    dplyr::group_by(visit_group) %>%
    dplyr::mutate(
      .neg_frac = slider::slide_index_dbl(
        .neg_ind, DT_round, ~ mean(.x, na.rm = TRUE),
        .before = lubridate::hours(pars$burial_window_hr))
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      spike_flag_v3 = dplyr::case_when(.signal ==  1 ~ "positive spike",
                                       .signal == -1 ~ "negative spike",
                                       TRUE ~ NA_character_),
      .burial_persist = dplyr::coalesce(.neg_frac >= pars$burial_min_frac, FALSE),
      flag_v3 = dplyr::case_when(
        .burial_persist ~ "Possible burial",
        .signal != 0    ~ spike_flag_v3,
        .rate_bad       ~ "high rate",
        TRUE ~ NA_character_)
    )
}
```

The stage-2 slope threshold comes from the seasonal-threshold archive. The
newest file is a 2026 one; we reuse it for the older years too (the sites are
stable, and a missing site just falls back to the 1 mg/L per step default).

```{r v3-run, cache = TRUE}
seasonal_thresholds <- read_parquet(here(
  "data", "derived", "auto_qaqc_files", "thresholds",
  "seasonal_thresholds20260304-T175010Z.parquet")) %>%
  filter(parameter == "DO") %>%
  distinct(site, season, .keep_all = TRUE)

# roughly 5 s per site-year, so this chunk is cached
DO_v3 <- do_split %>%
  map(~ map(.x, ~ generate_summary_statistics(.x) %>%
                  find_do_noise_v3(seasonal_thresholds))) %>%
  list_flatten() %>%
  list_rbind()
```

## 3-Way Method Comparison

```{r join-3way}
DO_three <- DO_both %>%
  left_join(
    DO_v3 %>% select(site, DT_round, flag_v3,
                     v3_burial  = .burial_persist,
                     v3_signal  = .signal,
                     v3_band_lo = .band_lo,
                     v3_band_hi = .band_hi),
    by = c("site", "DT_round")
  ) %>%
  mutate(
    m_old = old_burial,
    m_new = new_burial,
    m_v3  = coalesce(v3_burial, FALSE),
    # v3's non-burial output. Only a negative spike counts as a "v3 tag";
    # positive spikes and high-rate flags are ignored (-> "none" / grey).
    v3_tag = coalesce(flag_v3 == "negative spike", FALSE),
    # combo3: which methods flagged the point. v3 contributes one of three
    # states -- burial ("v3"), a tag ("v3 tag"), or nothing -- and a v3 tag
    # shows even when original or updated also flagged a burial there.
    .code = paste0(
      if_else(m_old, "O", ""), if_else(m_new, "U", ""),
      case_when(m_v3 ~ "B", v3_tag ~ "T", TRUE ~ "")
    ),
    .code = if_else(.code == "", "none", .code),
    combo3 = factor(
      recode(.code,
        "T"   = "v3 tag",
        "O"   = "original",       "U"   = "updated",
        "OU"  = "original+updated",
        "OT"  = "original+v3 tag", "UT"  = "updated+v3 tag",
        "OUT" = "original+updated+v3 tag",
        "B"   = "v3",
        "OB"  = "original+v3",     "UB"  = "updated+v3",
        "OUB" = "all three"),
      levels = c("none", "v3 tag",
                 "original", "updated", "original+updated",
                 "original+v3 tag", "updated+v3 tag", "original+updated+v3 tag",
                 "v3", "original+v3", "updated+v3", "all three")
    )
  ) %>%
  select(-.code)
```

Every confirmed-burial event in the data: contiguous runs of `user_burial`,
merged across gaps shorter than 24 h. `hit_*` is the fraction of the event's
points that method flagged "Possible burial" -- a rough per-event recall. Use
`plot_burial_3way(site, year)` to look at any of them.

```{r burial-events}
burial_events <- DO_three %>%
  filter(user_burial) %>%
  arrange(site, DT_round) %>%
  group_by(site) %>%
  mutate(event = cumsum(c(TRUE, diff(as.numeric(DT_round)) > 24 * 3600))) %>%
  group_by(site, event) %>%
  summarise(
    year  = min(year),
    start = min(DT_round),
    end   = max(DT_round),
    days  = round(as.numeric(difftime(max(DT_round), min(DT_round),
                                      units = "days")), 1),
    n_pts = n(),
    hit_original = round(mean(old_burial), 2),
    hit_updated  = round(mean(new_burial), 2),
    hit_v3       = round(mean(m_v3), 2),
    .groups = "drop"
  ) %>%
  select(site, year, start, end, days, n_pts,
         hit_original, hit_updated, hit_v3) %>%
  arrange(site, start)

burial_events
```

```{r plot-fn-3way}
# Same idea as plot_burial() but for all three methods. Points are colored by
# which methods flagged the point (combo3). A "v3 tag" (v3 negative spike)
# shows even when original or updated also called a burial there; the "+v3 tag"
# colors are the lighter shades. The grey lines are v3's stage-3
# band (rolling mean +/- half-width) -- what v3 compares each point against.
plot_burial_3way <- function(site_name, yr, data = DO_three) {

  d <- data %>% filter(site == site_name, year == yr) %>% arrange(DT_round)
  if (nrow(d) == 0) stop("no data for ", site_name, " ", yr)

  gt <- d %>% filter(user_burial)
  shapes <- if (nrow(gt) == 0) list() else {
    gt %>%
      mutate(run = cumsum(c(TRUE, diff(as.numeric(DT_round)) > 1200))) %>%
      group_by(run) %>%
      summarise(x0 = min(DT_round), x1 = max(DT_round), .groups = "drop") %>%
      purrr::pmap(function(run, x0, x1) list(
        type = "rect", xref = "x", yref = "paper",
        x0 = as.numeric(x0), x1 = as.numeric(x1), y0 = 0, y1 = 1,
        fillcolor = "orange", opacity = 0.15, line = list(width = 0)))
  }

  combo_cols <- c(
    "none"                     = "grey85",
    "v3 tag"                   = "#e6ab02",   # gold
    "original"                 = "#e41a1c",
    "updated"                  = "#377eb8",
    "original+updated"         = "#984ea3",
    "original+v3 tag"          = "#fb9a99",   # light red
    "updated+v3 tag"           = "#a6cee3",   # light blue
    "original+updated+v3 tag"  = "#cab2d6",   # light purple
    "v3"                       = "#4daf4a",
    "original+v3"              = "#ff7f00",
    "updated+v3"               = "#00ced1",
    "all three"                = "#f781bf"
  )

  band <- d %>%
    filter(!is.na(v3_band_lo)) %>%
    select(DT_round, visit_group, v3_band_lo, v3_band_hi) %>%
    pivot_longer(c(v3_band_lo, v3_band_hi), values_to = "y")

  p <- ggplot(d, aes(DT_round, mean)) + 
    geom_line(data = band,
              aes(DT_round, y, group = interaction(visit_group, name)),
              color = "grey70", linewidth = 0.3, inherit.aes = FALSE) +
    geom_point(aes(color = combo3,
                   text = paste0(format(DT_round), "<br>DO: ", round(mean, 2),
                                 "<br>", combo3)), size = 1) +
    geom_vline(data = filter(d, site_visit), aes(xintercept = DT_round),
               color = "red", linetype = "dashed", inherit.aes = FALSE) +
    scale_color_manual(values = combo_cols, drop = FALSE) +
    labs(title = paste(site_name, yr, "-- 3-way"),
         x = NULL, y = "DO (mg/L)", color = NULL) +
    theme_minimal()

  gp <- plotly::ggplotly(p, tooltip = "text")
  gp$x$layout$shapes <- shapes
  gp
}
```

```{r}
plot_burial_3way("bellvue", 2023)
```

```{r}
plot_burial_3way("archery", 2026)
```

## Confusion matrix

Point-level, against `user_burial` (technician-verified burial). A point is a
positive if the method flagged "Possible burial" there. 2026 is left out
because it has not been through manual verification yet, so there is no real
ground truth for it. `user_burial` is still imperfect in both directions for
2023-2025, so read these as rough. Burial is a small fraction of points, so
accuracy is not useful -- precision, recall, and specificity are.

```{r confusion-plot, fig.width = 9, fig.height = 3.6}
# 2x2 cell counts for one prediction vs user_burial
confusion_cells <- function(pred, actual) {
  pred   <- coalesce(pred, FALSE)
  actual <- coalesce(actual, FALSE)
  tibble(
    truth = if_else(actual, "burial", "no burial"),
    pred  = if_else(pred,   "flagged", "not flagged")
  ) %>% count(truth, pred, name = "n")
}

# 2023-2025 only -- 2026 is not verified
cm_data <- DO_three %>% filter(year <= 2025)

cm <- bind_rows(
  original = confusion_cells(cm_data$old_burial, cm_data$user_burial),
  updated  = confusion_cells(cm_data$new_burial, cm_data$user_burial),
  v3       = confusion_cells(cm_data$m_v3,        cm_data$user_burial),
  .id = "method"
) %>%
  mutate(
    truth = factor(truth, levels = c("burial", "no burial")),
    pred  = factor(pred,  levels = c("not flagged", "flagged")),
    cell  = case_when(
      truth == "burial"    & pred == "flagged"     ~ "TP",
      truth == "no burial" & pred == "flagged"     ~ "FP",
      truth == "burial"    & pred == "not flagged" ~ "FN",
      TRUE                                         ~ "TN"
    )
  )

# precision / recall per method, for the facet strip labels
cm_lab <- cm %>%
  group_by(method) %>%
  summarise(
    precision = sum(n[cell == "TP"]) / sum(n[cell %in% c("TP", "FP")]),
    recall    = sum(n[cell == "TP"]) / sum(n[cell %in% c("TP", "FN")]),
    .groups = "drop"
  ) %>%
  mutate(strip = sprintf("%s\nprecision %.2f   recall %.2f",
                         method, precision, recall))

cm <- cm %>%
  left_join(cm_lab, by = "method") %>%
  mutate(strip = factor(strip, levels = cm_lab$strip[
    order(match(cm_lab$method, c("original", "updated", "v3")))]))

ggplot(cm, aes(pred, truth, fill = cell)) +
  geom_tile(color = "white", linewidth = 1.5) +
  geom_text(aes(label = paste0(cell, "\n", scales::comma(n))),
            size = 3.3, lineheight = 0.95) +
  facet_wrap(~ strip) +
  scale_fill_manual(values = c(TP = "#a1d99b", FP = "#fc9272",
                               FN = "#fdbe85", TN = "grey90"),
                    guide = "none") +
  labs(x = "prediction", y = "truth (user_burial)",
       title = "Confusion matrix by method, point-level, 2023-2025") +
  theme_minimal(base_size = 11) +
  theme(panel.grid = element_blank())
```

# Next Steps

- Immediate next steps would probably be to combine both the old and new methods
into one stronger method with more logic gates. Then add that back to ross.wq.tools.
